---
title: 'Chemical Mixtures and Accelerated Biological Aging in CKD: NHANES Analysis'
author: "Reproducible NHANES analysis By January"
date: "`r format(Sys.Date())`"
output:
  word_document:
    toc: true
    toc_depth: '3'
  pdf_document:
    toc: true
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(20260227)
```

# Study Objective

Primary question: are environmental chemical mixtures associated with accelerated biological aging (PhenoAge acceleration), and is this association stronger among individuals with CKD?

# Configuration

```{r config}
config <- list(
  cycles = c("2005-2006", "2007-2008", "2009-2010", "2011-2012", "2013-2014", "2015-2016", "2017-2018"),  # Exclude 1999-2000, 2001-2002 (no standard PFAS), 2003-2004 (not used); retained cycles for compatible weights and full biomarkers
  exposure_family = "pfas", # pfas, metals, both
  q = 4,
  min_age = 20,
  ckd_def = "egfr_or_acr", # egfr_or_acr, egfr_only, acr_only
  use_residualized_phenoage = TRUE,   # PhenoAgeAccel is always residualized; option reserved for future.
  # Primary analysis uses fixed 4-PFAS set with complete-case; strict (qgcomp-compatible subset) is sensitivity.
  require_complete_exposure_set = FALSE,  # Legacy: primary is always fixed-set complete-case; no longer changes behavior.
  min_nonmissing_exposure_fraction = 0.6,
  min_nonmissing_exposures_floor = 2,
  include_clinical_covariates = TRUE,    # Not wired to primary adjustment set; reserved for future sensitivity.
  # Weight is resolved from audit; PFAS workflow prioritizes environmental subsample WTS* weights.
  weight_type = "auto",  # Auto-resolve from audit in the exposure-complete sample (PFAS excludes WTINT2YR/WTMEC2YR).
  # Multiple imputation (MICE) for missing covariates.
  use_mice = TRUE,        # If TRUE, Model B + sensitivities + supplementary use MI-pooled estimates (Rubin rules).
  mice_m = 24,            # Number of imputations: m >= % missing on most-missing covariate (alcohol3 ~24% pre-imputation), per White, Royston & Wood (2011).
  mice_maxit = 20,        # Iterations per chain.
  mice_seed = 20260227,   # Reproducibility seed for mice.
  verbose = TRUE
)
config
```

# Packages

```{r packages}
required_pkgs <- c(
  "nhanesA", "dplyr", "tidyr", "purrr", "stringr", "survey", "splines",
  "ggplot2", "broom", "knitr", "qgcomp", "forcats", "tibble",
  "mice", "mitools", "patchwork", "hexbin"
)
invisible(lapply(required_pkgs, function(p) {
  if (!requireNamespace(p, quietly = TRUE)) stop("Package not installed: ", p)
}))

library(nhanesA)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(survey)
options(survey.lonely.psu = "adjust")
library(splines)
library(ggplot2)
library(broom)
library(knitr)
library(qgcomp)
library(forcats)
library(tibble)
library(mice)
library(mitools)
library(patchwork)
```

# Helper Functions

```{r helpers}
cycle_to_suffix <- function(cycle) {
  map <- c(
    "1999-2000" = "A", "2001-2002" = "B", "2003-2004" = "C", "2005-2006" = "D",
    "2007-2008" = "E", "2009-2010" = "F", "2011-2012" = "G", "2013-2014" = "H",
    "2015-2016" = "I", "2017-2018" = "J", "2019-2020" = "K"
  )
  out <- unname(map[cycle])
  if (is.na(out)) stop("Cycle not mapped: ", cycle)
  out
}

safe_nhanes <- function(table_name, verbose = TRUE) {
  out <- tryCatch(nhanesA::nhanes(table_name), error = function(e) NULL)
  if (is.null(out) && verbose) message("Unavailable table: ", table_name)
  out
}

resolve_table <- function(cycle, candidates, domain = NULL, verbose = TRUE) {
  suf <- cycle_to_suffix(cycle)
  to_try <- paste0(candidates, "_", suf)
  for (tb in to_try) {
    dat <- safe_nhanes(tb, verbose = FALSE)
    if (!is.null(dat)) {
      if (verbose) message("Using table ", tb, " for cycle ", cycle)
      return(list(code = candidates[1], table = tb, data = dat))
    }
  }
  NULL
}

coalesce_num <- function(df, vars) {
  vars <- vars[vars %in% names(df)]
  if (length(vars) == 0) return(rep(NA_real_, nrow(df)))
  v <- lapply(df[vars], function(x) suppressWarnings(as.numeric(as.character(x))))
  dplyr::coalesce(!!!v)
}

parse_yes_no <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  x_chr <- toupper(trimws(as.character(x)))
  case_when(
    x_num == 1 ~ 1,
    x_num == 2 ~ 0,
    str_detect(x_chr, "^YES$") ~ 1,
    str_detect(x_chr, "^NO$") ~ 0,
    TRUE ~ NA_real_
  )
}

parse_smq040 <- function(x) {
  x_chr <- toupper(trimws(as.character(x)))
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  case_when(
    x_num %in% c(1, 2) ~ 1,
    x_num == 3 ~ 0,
    str_detect(x_chr, "EVERY DAY|SOME DAYS") ~ 1,
    str_detect(x_chr, "^NOT AT ALL$") ~ 0,
    TRUE ~ NA_real_
  )
}

# SMQ020 = "Smoked at least 100 cigarettes in entire life?" (the gate question for SMQ040).
# 1 = Yes (lifetime smoker), 2 = No (never smoker), 7/9 = Refused/Don't know -> NA.
parse_smq020 <- function(x) {
  x_chr <- toupper(trimws(as.character(x)))
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  case_when(
    x_num == 1 ~ 1,
    x_num == 2 ~ 0,
    str_detect(x_chr, "^YES$") ~ 1,
    str_detect(x_chr, "^NO$")  ~ 0,
    TRUE ~ NA_real_
  )
}

# Lifetime "ever drinker" indicator harmonized across NHANES cycles.
#   ALQ101 (1999-2016) = ">=12 alcoholic drinks in any one year ever?"  (1=Yes / 2=No)
#   ALQ110 (1999-2016) = ">=12 drinks in entire life?"  asked when ALQ101=No (1=Yes / 2=No)
#   ALQ111 (2017+)     = "Ever had a drink of any kind of alcohol?"     (1=Yes / 2=No)
# Returns 1 if ANY indicator is Yes, 0 if all available indicators are No, NA otherwise
# (refused / don't know / not asked).
parse_alq_ever <- function(alq101, alq110, alq111) {
  e101 <- parse_yes_no(alq101)
  e110 <- parse_yes_no(alq110)
  e111 <- parse_yes_no(alq111)
  case_when(
    !is.na(e111) & e111 == 1 ~ 1,
    !is.na(e101) & e101 == 1 ~ 1,
    !is.na(e110) & e110 == 1 ~ 1,
    !is.na(e111) & e111 == 0 ~ 0,
    !is.na(e101) & e101 == 0 & (is.na(e110) | e110 == 0) ~ 0,
    TRUE ~ NA_real_
  )
}

# "Current drinker" (drank in past 12 months) harmonized across cycles.
#   ALQ121 (2013-2018) = "Past 12 mo, how often drink alcoholic beverages?"
#                         0 = Never in last year, 1-10 = various frequencies, 77/99 = Refused/DK.
#   ALQ120Q (1999-2016) = "How often drink alcoholic beverages?" (numeric # of times)
#                         in the unit specified by ALQ120U (week/month/year).
#                         0 = Never; 1-365 = numeric answer; 777/999 = Refused/DK.
# Returns 1 if any-frequency drinker, 0 if explicit "never in past year", NA otherwise.
parse_alq_current_yr <- function(alq121, alq120q) {
  c121 <- suppressWarnings(as.numeric(as.character(alq121)))
  c120 <- suppressWarnings(as.numeric(as.character(alq120q)))
  case_when(
    !is.na(c121) & c121 %in% 1:10 ~ 1,
    !is.na(c121) & c121 == 0 ~ 0,
    !is.na(c120) & c120 >= 1 & c120 <= 365 ~ 1,
    !is.na(c120) & c120 == 0 ~ 0,
    TRUE ~ NA_real_
  )
}

parse_phq_item <- function(x) {
  x_chr <- toupper(trimws(as.character(x)))
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  case_when(
    x_num == 1 ~ 0, x_num == 2 ~ 1, x_num == 3 ~ 2, x_num == 4 ~ 3,
    x_num %in% 0:3 ~ x_num,
    str_detect(x_chr, "^NOT AT ALL$") ~ 0,
    str_detect(x_chr, "^SEVERAL DAYS$") ~ 1,
    str_detect(x_chr, "^MORE THAN HALF THE DAYS$") ~ 2,
    str_detect(x_chr, "^NEARLY EVERY DAY$") ~ 3,
    TRUE ~ NA_real_
  )
}

yesno01 <- function(x) {
  xc <- toupper(trimws(as.character(x)))
  xn <- suppressWarnings(as.numeric(as.character(x)))
  case_when(
    xn == 1 ~ 1,
    xn == 2 ~ 0,
    xc %in% c("YES", "1") ~ 1,
    xc %in% c("NO", "2") ~ 0,
    TRUE ~ NA_real_
  )
}

compute_egfr_2021 <- function(scr_mgdl, age, sex_raw) {
  sex_chr <- toupper(trimws(as.character(sex_raw)))
  sex_num <- suppressWarnings(as.numeric(as.character(sex_raw)))
  male <- case_when(
    sex_num == 1 ~ TRUE,
    sex_num == 2 ~ FALSE,
    sex_chr == "MALE" ~ TRUE,
    sex_chr == "FEMALE" ~ FALSE,
    TRUE ~ NA
  )
  kappa <- ifelse(male, 0.9, 0.7)
  alpha <- ifelse(male, -0.302, -0.241)
  sex_mult <- ifelse(male, 1, 1.012)
  142 * pmin(scr_mgdl / kappa, 1)^alpha * pmax(scr_mgdl / kappa, 1)^(-1.200) * (0.9938^age) * sex_mult
}

compute_acr <- function(ua, ucr, acr_direct = NULL) {
  calc <- ifelse(!is.na(ua) & !is.na(ucr) & ucr > 0, (ua / ucr) * 100, NA_real_)
  dplyr::coalesce(acr_direct, calc)
}

compute_phenoage_components <- function(df, include_scr = TRUE, include_alb = TRUE) {
  # Levine et al. (2018) coefficients expect: albumin g/L, creatinine µmol/L, glucose mmol/L.
  # NHANES provides: albumin g/dL, creatinine mg/dL, glucose often mg/dL (we standardize to mg/dL in pipeline).
  alb_gL   <- df$alb * 10
  scr_umol <- df$scr * 88.4
  glu_mmol <- df$glu / 18.018
  alb_term <- if (include_alb) (-0.0336 * alb_gL) else 0
  scr_term <- if (include_scr) ( 0.0095 * scr_umol) else 0
  glu_term <- 0.1953 * glu_mmol

  xb <- -19.907 +
    alb_term +
    scr_term +
    glu_term +
    0.0954 * log(pmax(df$crp, 1e-6)) -
    0.0120 * df$lym +
    0.0268 * df$mcv +
    0.3306 * df$rdw +
    0.00188 * df$alp +
    0.0554 * df$wbc +
    0.0804 * df$age

  xb <- pmin(pmax(xb, -50), 50)
  gamma_const <- (exp(120 * 0.0076927) - 1) / 0.0076927
  hazard <- exp(xb) * gamma_const
  mort <- -expm1(-hazard)
  mort <- pmin(pmax(mort, 1e-12), 1 - 1e-12)
  pheno <- 141.50 + log(-0.00553 * log1p(-mort)) / 0.090165

  list(xb = xb, mort = mort, phenoage = pheno)
}

safe_phenoage <- function(df) {
  compute_phenoage_components(df, include_scr = TRUE, include_alb = TRUE)$phenoage
}

safe_qgcomp_extract <- function(fit_obj, model_label, n_obs) {
  if (is.null(fit_obj)) {
    return(tibble(model = model_label, beta = NA_real_, conf.low = NA_real_, conf.high = NA_real_, p.value = NA_real_, N = n_obs))
  }
  cf <- coef(fit_obj)
  psi_name <- grep("^psi", names(cf), value = TRUE)[1]
  if (is.na(psi_name)) psi_name <- grep("psi", names(cf), value = TRUE)[1]
  if (is.na(psi_name)) {
    message("qgcomp terms for [", model_label, "]: no psi found in ", paste(names(cf), collapse = ", "))
    return(tibble(model = model_label, beta = NA_real_, conf.low = NA_real_, conf.high = NA_real_, p.value = NA_real_, N = n_obs))
  }
  beta <- unname(cf[psi_name])
  vc <- tryCatch(vcov(fit_obj), error = function(e) NULL)
  if (!is.null(vc) && psi_name %in% rownames(vc) && psi_name %in% colnames(vc)) {
    se <- sqrt(vc[psi_name, psi_name])
    lo <- beta - 1.96 * se
    hi <- beta + 1.96 * se
    p_val <- 2 * pnorm(-abs(beta / se))
  } else {
    lo <- hi <- p_val <- NA_real_
  }
  tibble(model = model_label, beta = beta, conf.low = lo, conf.high = hi, p.value = p_val, N = n_obs)
}

resolve_weight_var <- function(df, weight_type = "auto", exposure_family = "pfas") {
  cols <- names(df)
  if (weight_type != "auto" && weight_type != "MEC") {
    if (!(weight_type %in% cols)) stop("Requested weight variable not found: ", weight_type)
    return(list(var = weight_type, note = paste("User-specified weight:", weight_type)))
  }
  if ("WTMEC2YR" %in% cols) return(list(var = "WTMEC2YR", note = "Generic fallback: using MEC weight WTMEC2YR because no exposure-specific subsample weight has been resolved. In PFAS analyses, final design should normally use the PFAS-specific subsample weight identified by the audit/auto-selection logic."))
  stop("No supported NHANES weight variable found.")
}

apply_lod_rule <- function(df, varname) {
  x <- suppressWarnings(as.numeric(as.character(df[[varname]])))

  base <- sub("^(LBX|URX)", "", varname)
  lc_candidates <- c(
    paste0(varname, "LC"),
    paste0("LBD", base, "LC"),
    paste0("URD", base, "LC"),
    # NHANES often uses *L (not *LC) as the detection-limit flag.
    paste0(varname, "L"),
    paste0("LBD", base, "L"),
    paste0("URD", base, "L")
  )
  lc_candidates <- lc_candidates[lc_candidates %in% names(df)]

  if (length(lc_candidates) > 0) {
    lc <- df[[lc_candidates[1]]]
    lc_num <- suppressWarnings(as.numeric(as.character(lc)))
    lc_chr <- toupper(trimws(as.character(lc)))
    below <- (!is.na(lc_num) & lc_num == 1) | str_detect(lc_chr, "BELOW|<")
    below[is.na(below)] <- FALSE
    # NHANES convention: when LC==1, the concentration column typically holds the LOD value.
    # Standard replacement: value / sqrt(2).
    x[below] <- x[below] / sqrt(2)
    # If value is still NA/non-positive after that, use smallest observed LOD / sqrt(2).
    lod_vals <- x[below & x > 0]
    fallback <- if (length(lod_vals) > 0) min(lod_vals, na.rm = TRUE) else NA_real_
    x[below & (is.na(x) | x <= 0)] <- fallback
    return(x)
  }
  # If no LOD indicator exists, do not impute missing values.
  x
}
```

# Data Acquisition and Cycle Availability

```{r cycle-download}
# Cycle-aware: 2003-2004 (C) uses Lxx_ names (L40, L25, L06CRP, L06ALB, L24PFC, L06BMT, L06UHM); 2005+ use BIOPRO/CBC/CRP/PFC; 2013+ PFAS; I/J use HSCRP.
table_candidates <- list(
  DEMO = c("DEMO"),
  BMX = c("BMX"),
  SMQ = c("SMQ"),
  ALQ = c("ALQ"),
  PAQ = c("PAQ"),
  DIQ = c("DIQ"),
  BPQ = c("BPQ"),
  MCQ = c("MCQ"),
  BIOPRO = c("L40", "BIOPRO"),
  CBC = c("L25", "CBC"),
  CRP = c("L06CRP", "CRP", "HSCRP"),
  ALB_CR = c("L06ALB", "ALB_CR"),
  GLU = c("L10", "GLU"),
  PFAS = c("L24PFC", "PFAS", "PFC", "P_PFAS"),
  BLOOD_METALS = c("L06BMT", "PBCD"),
  URINE_METALS = c("L06UHM", "UHM", "UM")
)

download_cycle <- function(cycle, exposure_family = "pfas", verbose = TRUE) {
  core_domains <- c("DEMO", "BMX", "SMQ", "ALQ", "PAQ", "DIQ", "BPQ", "MCQ", "BIOPRO", "CBC", "CRP", "ALB_CR", "GLU")
  exp_domains <- switch(
    exposure_family,
    pfas = c("PFAS"),
    metals = c("BLOOD_METALS", "URINE_METALS"),
    both = c("PFAS", "BLOOD_METALS", "URINE_METALS"),
    c("PFAS")
  )
  all_domains <- c(core_domains, exp_domains)

  resolved <- lapply(all_domains, function(dom) resolve_table(cycle, table_candidates[[dom]], domain = dom, verbose = verbose))
  names(resolved) <- all_domains
  resolved <- resolved[!vapply(resolved, is.null, logical(1))]
  if (!"DEMO" %in% names(resolved)) return(NULL)

  dat <- resolved$DEMO$data
  if (!"SEQN" %in% names(dat)) return(NULL)

  for (nm in setdiff(names(resolved), "DEMO")) {
    d <- resolved[[nm]]$data
    # Rename glucose vars from GLU table before joining so they do not collide with BIOPRO/L40.
    if (nm == "GLU") {
      if ("LBXGLU" %in% names(d)) names(d)[names(d) == "LBXGLU"] <- "LBXGLU_GLU"
      if ("LBXSGL" %in% names(d)) names(d)[names(d) == "LBXSGL"] <- "LBXSGL_GLU"
    }
    if ("SEQN" %in% names(d)) dat <- left_join(dat, d, by = "SEQN")
  }
  # After merge: define explicit glucose sources. Primary analysis uses MEC chemistry glucose for measurement consistency in PhenoAge construction.
  dat$glu_mec  <- coalesce_num(dat, c("LBXGLU", "LBXSGL"))
  dat$glu_glu  <- coalesce_num(dat, c("LBXGLU_GLU", "LBXSGL_GLU"))
  dat$glu     <- dat$glu_mec
  dat$cycle <- cycle
  # Store provenance for key tables so we can audit weight/source compatibility.
  if ("CRP" %in% names(resolved)) dat$crp_table <- resolved$CRP$table
  if ("GLU" %in% names(resolved)) dat$glu_table <- resolved$GLU$table
  # Warn if joins created .x/.y duplicate columns (would break coalesce_num).
  dup_suffix_cols <- names(dat)[grepl("\\.x$|\\.y$", names(dat))]
  if (length(dup_suffix_cols) > 0) {
    warning("Cycle ", cycle, ": duplicate columns from join: ", paste(dup_suffix_cols, collapse = ", "))
  }
  dat
}

cycle_data <- lapply(config$cycles, download_cycle, exposure_family = config$exposure_family, verbose = config$verbose)
cycle_data <- cycle_data[!vapply(cycle_data, is.null, logical(1))]
if (length(cycle_data) == 0) stop("No cycles were successfully downloaded.")

all_cols <- sort(unique(unlist(lapply(cycle_data, names))))
cycle_data <- lapply(cycle_data, function(df) {
  miss <- setdiff(all_cols, names(df))
  if (length(miss) > 0) for (m in miss) df[[m]] <- NA
  df <- df[, all_cols, drop = FALSE]
  df %>% mutate(across(everything(), as.character))
})
nhanes_raw <- bind_rows(cycle_data)

kable(
  data.frame(
    cycle = names(table(config$cycles))[names(table(config$cycles)) %in% unique(nhanes_raw$cycle)],
    n = as.integer(table(nhanes_raw$cycle))
  ),
  caption = "Cycle availability and downloaded participant counts"
)
```

# Variable Construction: PhenoAge, CKD, Covariates

```{r construct-vars}
# Defensive guard: ensure cycle-dependent NHANES columns exist before referencing them
# in the mutate below. cycle_data harmonization (Data Acquisition section) already pads
# cycles to the union of column names, so these will normally be present -- but if the
# user reconfigures `config$cycles` to drop certain ranges, some columns can vanish from
# the union (e.g., ALQ142 only exists in 2017+; PAD200/PAD320 only exist in 1999-2006).
# This block guarantees they exist as NA so the harmonized derivations below still run.
required_optional_cols <- c(
  # Smoking
  "SMQ020", "SMQ040",
  # Alcohol (cycle-dependent question wording -- see parser comments above)
  "ALQ101", "ALQ110", "ALQ111", "ALQ120Q", "ALQ121", "ALQ142", "ALQ151",
  # Physical activity (GPAQ in 2007+, legacy module in 1999-2006)
  "PAD200", "PAD320", "PAQ650", "PAQ665"
)
for (.col in required_optional_cols) {
  if (!.col %in% names(nhanes_raw)) nhanes_raw[[.col]] <- NA_character_
}
rm(.col)

dpq_vars <- names(nhanes_raw)[str_detect(names(nhanes_raw), "^DPQ0")]

dat <- nhanes_raw %>%
  mutate(across(any_of(dpq_vars), parse_phq_item)) %>%
  mutate(
    age = suppressWarnings(as.numeric(as.character(RIDAGEYR))),
    sex = case_when(
      suppressWarnings(as.numeric(as.character(RIAGENDR))) == 1 ~ "Male",
      suppressWarnings(as.numeric(as.character(RIAGENDR))) == 2 ~ "Female",
      toupper(trimws(as.character(RIAGENDR))) == "MALE" ~ "Male",
      toupper(trimws(as.character(RIAGENDR))) == "FEMALE" ~ "Female",
      TRUE ~ NA_character_
    ),
    race_eth = {
      r3 <- suppressWarnings(as.numeric(as.character(RIDRETH3)))
      r1 <- suppressWarnings(as.numeric(as.character(RIDRETH1)))
      r3_chr <- toupper(trimws(as.character(RIDRETH3)))
      r1_chr <- toupper(trimws(as.character(RIDRETH1)))
      case_when(
        !is.na(r3) & r3 %in% c(1, 2) ~ "Hispanic",
        !is.na(r3) & r3 == 3 ~ "Non-Hispanic White",
        !is.na(r3) & r3 == 4 ~ "Non-Hispanic Black",
        !is.na(r3) & r3 == 6 ~ "Non-Hispanic Asian",
        !is.na(r3) & r3 %in% c(5, 7) ~ "Other/Multi",
        is.na(r3) & r1 %in% c(1, 2) ~ "Hispanic",
        is.na(r3) & r1 == 3 ~ "Non-Hispanic White",
        is.na(r3) & r1 == 4 ~ "Non-Hispanic Black",
        is.na(r3) & r1 == 5 ~ "Other/Multi",
        str_detect(r3_chr, "NON-HISPANIC WHITE") ~ "Non-Hispanic White",
        str_detect(r3_chr, "NON-HISPANIC BLACK") ~ "Non-Hispanic Black",
        str_detect(r3_chr, "NON-HISPANIC ASIAN") ~ "Non-Hispanic Asian",
        str_detect(r3_chr, "OTHER RACE|MULTI") ~ "Other/Multi",
        str_detect(r3_chr, "^MEXICAN AMERICAN$|^OTHER HISPANIC$|^HISPANIC$") ~ "Hispanic",
        str_detect(r1_chr, "NON-HISPANIC WHITE") ~ "Non-Hispanic White",
        str_detect(r1_chr, "NON-HISPANIC BLACK") ~ "Non-Hispanic Black",
        str_detect(r1_chr, "OTHER RACE|MULTI") ~ "Other/Multi",
        str_detect(r1_chr, "^MEXICAN AMERICAN$|^OTHER HISPANIC$|^HISPANIC$") ~ "Hispanic",
        TRUE ~ NA_character_
      )
    },
    education = {
      e_num <- suppressWarnings(as.numeric(as.character(DMDEDUC2)))
      e_chr <- toupper(trimws(as.character(DMDEDUC2)))
      case_when(
        e_num %in% c(1, 2) ~ "<HS",
        e_num == 3 ~ "HS/GED",
        e_num %in% c(4, 5) ~ ">HS",
        str_detect(e_chr, "LESS THAN 9TH|9TH.*12TH") ~ "<HS",
        str_detect(e_chr, "HIGH SCHOOL|GED") ~ "HS/GED",
        str_detect(e_chr, "SOME COLLEGE|COLLEGE GRADUATE") ~ ">HS",
        TRUE ~ NA_character_
      )
    },
    pir = suppressWarnings(as.numeric(as.character(INDFMPIR))),
    bmi = suppressWarnings(as.numeric(as.character(BMXBMI))),
    # ---- Categorical-variable derivations ---------------------------------
    # Smoking: NHANES uses a 2-stage instrument. SMQ020 ("Ever smoked >=100 cigarettes
    # in life?") is the GATE question; SMQ040 ("Do you now smoke?") is asked only of
    # SMQ020==Yes respondents. Therefore an SMQ040-only derivation silently drops every
    # never-smoker (SMQ020==No) to NA and excludes them from any complete-case modeling
    # sample. Below we derive the standard 3-level smoker3 (Never/Former/Current) using
    # SMQ020 + SMQ040, and a true binary smoker (Current vs. Not-current) that retains
    # never-smokers as 0. This binary is what downstream models continue to use.
    smoker3 = {
      ever <- parse_smq020(SMQ020)
      smq040_chr <- toupper(trimws(as.character(SMQ040)))
      smq040_num <- suppressWarnings(as.numeric(as.character(SMQ040)))
      curr <- case_when(
        smq040_num %in% c(1, 2) ~ 1,
        smq040_num == 3 ~ 0,
        str_detect(smq040_chr, "EVERY DAY|SOME DAYS") ~ 1,
        str_detect(smq040_chr, "^NOT AT ALL$") ~ 0,
        TRUE ~ NA_real_
      )
      case_when(
        ever == 0 ~ "Never",
        ever == 1 & curr == 1 ~ "Current",
        ever == 1 & curr == 0 ~ "Former",
        TRUE ~ NA_character_
      )
    },
    smoker = case_when(
      smoker3 == "Current" ~ 1,
      smoker3 %in% c("Never", "Former") ~ 0,
      TRUE ~ NA_real_
    ),
    # Alcohol: harmonized 3-level Never/Former/Current drinker, parallel to smoker3.
    # The previous derivation used ALQ151 ("ever heavy drinker") only, which (a) silently
    # dropped never-drinkers (skip pattern), (b) returned NA for the entire 2017-2018 (J)
    # cycle (variable discontinued), and (c) measured a different construct than the
    # NHANES-literature standard. The new derivation uses the cycle-harmonized helpers:
    #   parse_alq_ever()        : lifetime ever-drinker from ALQ101 / ALQ110 / ALQ111
    #   parse_alq_current_yr()  : drank in past 12 mo from ALQ121 (2013+) or ALQ120Q (1999-2016)
    # Never-drinkers are now coded as "Never" (instead of NA), and J-cycle participants
    # get a defined value via ALQ111 + ALQ121.
    alcohol3 = {
      ever <- parse_alq_ever(ALQ101, ALQ110, ALQ111)
      curr <- parse_alq_current_yr(ALQ121, ALQ120Q)
      case_when(
        ever == 0 ~ "Never",
        ever == 1 & curr == 1 ~ "Current",
        ever == 1 & curr == 0 ~ "Former",
        TRUE ~ NA_character_
      )
    },
    # Binary alcohol covariate (used by downstream models): "Current drinker" vs.
    # "Not current drinker" (combines never + former). This matches the NHANES literature
    # convention and is far less sparse than the previous ALQ151-based binary.
    alcohol = case_when(
      alcohol3 == "Current" ~ 1,
      alcohol3 %in% c("Never", "Former") ~ 0,
      TRUE ~ NA_real_
    ),
    # Physical activity: harmonized across the 2006/2007 GPAQ transition.
    #   PAQ650 (vigorous recreational, 2007+) OR PAQ665 (moderate recreational, 2007+)
    #   PAD200 (vigorous past 30 days, 1999-2006) OR PAD320 (moderate past 30 days, 1999-2006)
    # Either questionnaire format counts as "physically active" for the harmonized binary;
    # otherwise the participant is asked one set or the other depending on cycle.
    phys_active = case_when(
      parse_yes_no(PAQ650) == 1 | parse_yes_no(PAQ665) == 1 ~ 1,
      parse_yes_no(PAQ650) == 0 & parse_yes_no(PAQ665) == 0 ~ 0,
      parse_yes_no(PAD200) == 1 | parse_yes_no(PAD320) == 1 ~ 1,
      parse_yes_no(PAD200) == 0 & parse_yes_no(PAD320) == 0 ~ 0,
      TRUE ~ NA_real_
    ),
    # Diabetes: DIQ010 ("Doctor told you have diabetes?") codes 1=Yes, 2=No, 3=Borderline,
    # 7/9=Refused/DK. yesno01() maps 1->1, 2->0, and ALL others (including Borderline) to NA.
    # This drops borderline-diabetics from the analytic sample. If PI prefers, borderline
    # can be coded as 1 (any diabetes) or 0 (definite diabetes only).
    diabetes = yesno01(DIQ010),
    # Hypertension: BPQ020 ("Doctor told you have HBP?") asked of all adults; no gate.
    hypertension = yesno01(BPQ020),
    # CVD composite: ANY of CHF (MCQ160B), CHD (MCQ160C), heart attack (MCQ160E), stroke
    # (MCQ160F). Not asked behind a gate; respondent NA in all four -> overall NA.
    cvd = case_when(
      parse_yes_no(MCQ160B) == 1 | parse_yes_no(MCQ160C) == 1 | parse_yes_no(MCQ160E) == 1 | parse_yes_no(MCQ160F) == 1 ~ 1,
      parse_yes_no(MCQ160B) == 0 & parse_yes_no(MCQ160C) == 0 & parse_yes_no(MCQ160E) == 0 & parse_yes_no(MCQ160F) == 0 ~ 0,
      TRUE ~ NA_real_
    ),
    phenoage_accel = NA_real_
  )

# Compute biomarkers, CKD markers, and design variables outside mutate (avoids cur_data_all()).
dat$alb <- coalesce_num(dat, c("LBXSAL"))
dat$scr <- coalesce_num(dat, c("LBXSCR", "LBDSCRSI"))
# Glucose: glu_mec / glu_glu / glu set in download_cycle (primary uses glu_mec by design).
if (!"glu" %in% names(dat)) dat$glu <- dat$glu_mec
dat$crp <- coalesce_num(dat, c("LBXCRP", "LBXHSCRP"))
dat$lym <- coalesce_num(dat, c("LBXLYPCT"))
dat$mcv <- coalesce_num(dat, c("LBXMCV", "LBXMCVSI"))
dat$rdw <- coalesce_num(dat, c("LBXRW", "LBXRDW"))
dat$alp <- coalesce_num(dat, c("LBXAPSI", "LBXSAPSI"))
dat$wbc <- coalesce_num(dat, c("LBXWBCSI", "LBXWBC"))

dat$ualb <- coalesce_num(dat, c("URXUMA"))
dat$ucr <- coalesce_num(dat, c("URXUCR", "URDUCR"))
dat$acr_direct <- coalesce_num(dat, c("URDACT"))
dat$acr <- compute_acr(dat$ualb, dat$ucr, dat$acr_direct)
dat$egfr <- compute_egfr_2021(dat$scr, dat$age, dat$RIAGENDR)
dat$ckd_egfr <- ifelse(!is.na(dat$egfr), as.integer(dat$egfr < 60), NA_integer_)
dat$ckd_acr <- ifelse(!is.na(dat$acr), as.integer(dat$acr >= 30), NA_integer_)
dat$ckd <- case_when(
  config$ckd_def == "egfr_or_acr" & (dat$ckd_egfr == 1 | dat$ckd_acr == 1) ~ 1L,
  config$ckd_def == "egfr_or_acr" & (!is.na(dat$ckd_egfr) | !is.na(dat$ckd_acr)) ~ 0L,
  config$ckd_def == "egfr_only" ~ dat$ckd_egfr,
  config$ckd_def == "acr_only" ~ dat$ckd_acr,
  TRUE ~ NA_integer_
)
dat$ckd_stage <- cut(
  dat$egfr, breaks = c(-Inf, 15, 30, 45, 60, 90, Inf),
  labels = c("G5", "G4", "G3b", "G3a", "G2", "G1"),
  right = FALSE  # KDIGO: left-closed intervals [lo, hi). E.g., 60 -> G2, 59.9 -> G3a.
)
dat$strata <- coalesce_num(dat, c("SDMVSTRA"))
dat$psu <- coalesce_num(dat, c("SDMVPSU"))

# Coding audits (raw -> derived) to prevent silent recode failures.
kable(
  dat %>% count(cycle, ridreth3_missing = is.na(RIDRETH3)) %>% arrange(cycle, ridreth3_missing),
  caption = "Race audit: availability of RIDRETH3 by cycle"
)
kable(as.data.frame(table(dat$RIDRETH1, useNA = "ifany")), caption = "Raw RIDRETH1 distribution")
kable(as.data.frame(table(dat$RIDRETH3, useNA = "ifany")), caption = "Raw RIDRETH3 distribution")
kable(as.data.frame(table(dat$race_eth, useNA = "ifany")), caption = "Derived race_eth distribution")

kable(as.data.frame(table(dat$SMQ040, useNA = "ifany")), caption = "Raw SMQ040 distribution")
kable(as.data.frame(table(dat$DIQ010, useNA = "ifany")), caption = "Raw DIQ010 distribution")
kable(as.data.frame(table(dat$BPQ020, useNA = "ifany")), caption = "Raw BPQ020 distribution")
kable(as.data.frame(table(dat$smoker, useNA = "ifany")), caption = "Derived smoker distribution")
kable(as.data.frame(table(dat$diabetes, useNA = "ifany")), caption = "Derived diabetes distribution")
kable(as.data.frame(table(dat$hypertension, useNA = "ifany")), caption = "Derived hypertension distribution")

# Codebook-aligned harmonization checks (raw coding support by cycle).
check_raw_levels <- function(df, var, allowed_patterns) {
  if (!(var %in% names(df))) return(tibble(variable = var, cycle = NA_character_, n_unexpected = NA_integer_, unexpected_examples = NA_character_))
  df %>%
    mutate(.raw = toupper(trimws(as.character(.data[[var]])))) %>%
    group_by(cycle) %>%
    summarise(
      n_nonmissing = sum(!is.na(.raw) & .raw != ""),
      n_unexpected = sum(!is.na(.raw) & .raw != "" & !str_detect(.raw, allowed_patterns)),
      unexpected_examples = paste(head(unique(.raw[!is.na(.raw) & .raw != "" & !str_detect(.raw, allowed_patterns)]), 6), collapse = ", "),
      .groups = "drop"
    ) %>%
    mutate(variable = var) %>%
    select(variable, cycle, n_nonmissing, n_unexpected, unexpected_examples)
}

harmonization_checks <- bind_rows(
  check_raw_levels(dat, "RIAGENDR", "^(1|2|MALE|FEMALE)$"),
  check_raw_levels(dat, "DMDEDUC2", "^(1|2|3|4|5|7|9)$|LESS THAN 9TH|9-11TH|HIGH SCHOOL|GED|SOME COLLEGE|COLLEGE GRADUATE|DON'T KNOW|DONT KNOW|REFUSED"),
  check_raw_levels(dat, "SMQ040", "^(1|2|3|7|9)$|EVERY DAY|SOME DAYS|NOT AT ALL|REFUSED|DON'T KNOW|DONT KNOW"),
  check_raw_levels(dat, "DIQ010", "^(1|2|3|7|9)$|YES|NO|BORDERLINE|REFUSED|DON'T KNOW|DONT KNOW"),
  check_raw_levels(dat, "BPQ020", "^(1|2|7|9)$|YES|NO|REFUSED|DON'T KNOW|DONT KNOW")
)
kable(harmonization_checks, caption = "Codebook-aligned raw coding checks by cycle. Non-zero unexpected counts indicate coding values not covered by harmonization rules.")

# --- CRP unit detection and conversion ---
# PhenoAge (Levine 2018) uses CRP in mg/dL. NHANES LBXCRP is typically mg/dL;
# LBXHSCRP or CRP tables in some cycles may report mg/L.
# Step 1: check variable/table provenance.
crp_by_cycle <- dat %>%
  group_by(cycle) %>%
  summarise(
    source_table = if ("crp_table" %in% names(pick(everything()))) paste(unique(na.omit(crp_table)), collapse = ",") else "unknown",
    n_obs = sum(!is.na(crp)),
    median_crp = median(crp, na.rm = TRUE),
    p75 = quantile(crp, 0.75, na.rm = TRUE),
    p90 = quantile(crp, 0.90, na.rm = TRUE),
    .groups = "drop"
  )
kable(crp_by_cycle, caption = "CRP diagnostic by cycle -- verify units (PhenoAge expects mg/dL). If source table is HSCRP, values may be mg/L.")

# Step 2: provenance-based detection (strongest signal).
crp_tables_used <- if ("crp_table" %in% names(dat)) unique(na.omit(dat$crp_table)) else character(0)
provenance_suggests_mgl <- length(crp_tables_used) > 0 && any(str_detect(crp_tables_used, regex("HSCRP", ignore_case = TRUE)))
# Step 3: distribution-based backup -- require median > 1.5 AND p75 > 3 to trigger conversion.
crp_overall_med <- median(dat$crp, na.rm = TRUE)
crp_overall_p75 <- quantile(dat$crp, 0.75, na.rm = TRUE)
distribution_suggests_mgl <- !is.na(crp_overall_med) && crp_overall_med > 1.5 &&
                             !is.na(crp_overall_p75) && crp_overall_p75 > 3

if (provenance_suggests_mgl || distribution_suggests_mgl) {
  reason <- if (provenance_suggests_mgl) {
    paste0("CRP source table(s) [", paste(crp_tables_used, collapse = ", "), "] suggest mg/L.")
  } else {
    paste0("Distribution heuristic: median=", round(crp_overall_med, 3), ", p75=", round(crp_overall_p75, 3), ".")
  }
  message(">>> Converting CRP from mg/L to mg/dL (/ 10). Reason: ", reason)
  dat$crp <- dat$crp / 10
} else {
  message(">>> CRP assumed to be in mg/dL (median=", round(crp_overall_med, 4),
          ", p75=", round(crp_overall_p75, 4), "). Verify against NHANES documentation.")
}

# --- Glucose provenance diagnostic ---
# Coerce glu to numeric (nhanes_raw has character columns from bind_rows); needed for summarise and later unit conversion.
dat$glu <- suppressWarnings(as.numeric(as.character(dat$glu)))
glu_by_cycle <- dat %>%
  group_by(cycle) %>%
  summarise(
    source_table = if ("glu_table" %in% names(pick(everything()))) paste(unique(na.omit(glu_table)), collapse = ",") else "unknown",
    n_glu = sum(!is.na(glu)),
    median_glu = median(glu, na.rm = TRUE),
    min_glu = min(glu, na.rm = TRUE),
    max_glu = max(glu, na.rm = TRUE),
    .groups = "drop"
  )
kable(glu_by_cycle, caption = "Glucose provenance by cycle. If source_table is a fasting component (GLU_*), use a compatible fasting subsample weight.")

glu_med <- median(dat$glu, na.rm = TRUE)
if (!is.na(glu_med) && glu_med < 6) {
  warning("Glucose median is < 6 -- this looks like mmol/L. Converting to mg/dL ( x 18.018).")
  dat$glu <- dat$glu * 18.018
}

# Warn based on stored table provenance, not just weight variable existence.
glu_tables_used <- if ("glu_table" %in% names(dat)) unique(na.omit(dat$glu_table)) else character(0)
glu_is_fasting <- length(glu_tables_used) > 0 && any(str_detect(glu_tables_used, regex("^GLU_", ignore_case = TRUE)))
if (glu_is_fasting) {
  message(">>> IMPORTANT: Glucose sourced from fasting component table(s) [",
          paste(glu_tables_used, collapse = ", "),
          "]. A compatible fasting subsample weight (e.g., WTSAF2YR) is required. ",
          "Using a non-fasting or otherwise incompatible weight can bias estimates. ",
          "Review the weight audit; override config$weight_type only if auto-selection does not resolve the correct subsample weight.")
} else if (length(glu_tables_used) > 0) {
  message(">>> Glucose source table(s): ", paste(glu_tables_used, collapse = ", "),
          ". Verify whether this is a fasting component.")
}

# ------------------------------------------------------------
# CLEAN SPECIAL NHANES MISSING CODES FOR LAB VARIABLES
# ------------------------------------------------------------
clean_nhanes_numeric <- function(x) {
  x <- suppressWarnings(as.numeric(as.character(x)))
  special_codes <- c(
    7, 77, 777, 7777, 77777,
    9, 99, 999, 9999, 99999,
    888, 8888, 88888
  )
  x[x %in% special_codes] <- NA_real_
  x
}

dat$alb <- clean_nhanes_numeric(dat$alb)
dat$scr <- clean_nhanes_numeric(dat$scr)
dat$glu <- clean_nhanes_numeric(dat$glu)
dat$crp <- clean_nhanes_numeric(dat$crp)
dat$lym <- clean_nhanes_numeric(dat$lym)
dat$mcv <- clean_nhanes_numeric(dat$mcv)
dat$rdw <- clean_nhanes_numeric(dat$rdw)
dat$alp <- clean_nhanes_numeric(dat$alp)
dat$wbc <- clean_nhanes_numeric(dat$wbc)

# ------------------------------------------------------------
# HARD PLAUSIBILITY CHECK (retain CRP >1 mg/dL in primary; flag as acute)
# ------------------------------------------------------------
dat <- dat %>%
  mutate(
    crp_acute = ifelse(!is.na(crp) & crp > 1, 1L, 0L),
    wbc = ifelse(wbc > 50, NA_real_, wbc),
    glu = ifelse(glu > 600, NA_real_, glu),
    alp = ifelse(alp > 500, NA_real_, alp)
  )

# --- General biomarker plausibility ---
biomarker_ranges <- list(
  alb  = c(1, 7,    "g/dL"),
  scr  = c(0.1, 25, "mg/dL"),
  glu  = c(20, 600,  "mg/dL"),
  crp  = c(0.001, 3, "mg/dL (post-conversion if applied)"),
  lym  = c(1, 80,    "%"),
  mcv  = c(50, 150,  "fL"),
  rdw  = c(8, 30,    "%"),
  alp  = c(5, 500,   "U/L"),
  wbc  = c(1, 50,    "1000 cells/uL")
)
biomarker_check <- lapply(names(biomarker_ranges), function(nm) {
  vals <- dat[[nm]]
  rng  <- biomarker_ranges[[nm]]
  lo   <- as.numeric(rng[1]); hi <- as.numeric(rng[2]); unit <- rng[3]
  v    <- vals[!is.na(vals)]
  n_out <- sum(v < lo | v > hi)
  tibble(biomarker = nm, unit = unit, n_obs = length(v),
         min = if (length(v) > 0) min(v) else NA_real_,
         p25 = if (length(v) > 0) quantile(v, 0.25) else NA_real_,
         median = if (length(v) > 0) median(v) else NA_real_,
         p75 = if (length(v) > 0) quantile(v, 0.75) else NA_real_,
         max = if (length(v) > 0) max(v) else NA_real_,
         expected_lo = lo, expected_hi = hi,
         n_out_of_range = n_out)
}) %>% bind_rows()
kable(biomarker_check, caption = "PhenoAge biomarker ranges after unit conversion (verify units)")

for (nm in names(biomarker_ranges)) {
  v <- dat[[nm]][!is.na(dat[[nm]])]
  lo <- as.numeric(biomarker_ranges[[nm]][1])
  hi <- as.numeric(biomarker_ranges[[nm]][2])
  pct_out <- if (length(v) > 0) mean(v < lo | v > hi) else 0
  if (pct_out > 0.20) warning(sprintf(
    "Biomarker '%s': %.1f%% of values outside plausible range [%s, %s]. Check units.",
    nm, pct_out * 100, lo, hi
  ))
}

# Recompute PhenoAge after any unit conversions (CRP mg/L->mg/dL, glucose mmol/L->mg/dL).
pheno_core <- compute_phenoage_components(dat, include_scr = TRUE, include_alb = TRUE)
dat$phenoage <- pheno_core$phenoage

message(
  "PhenoAge diagnostics -- xb median=", round(stats::median(pheno_core$xb, na.rm = TRUE), 3),
  ", mort median=", format(stats::median(pheno_core$mort, na.rm = TRUE), scientific = TRUE, digits = 3),
  ", pct(mort>0.9999)=", round(mean(pheno_core$mort > 0.9999, na.rm = TRUE) * 100, 1), "%",
  ", PhenoAge SD=", round(stats::sd(dat$phenoage, na.rm = TRUE), 3)
)
ph_q <- stats::quantile(dat$phenoage, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
message(
  "PhenoAge five-number summary -- min=", round(ph_q[[1]], 3),
  ", p25=", round(ph_q[[2]], 3),
  ", median=", round(ph_q[[3]], 3),
  ", p75=", round(ph_q[[4]], 3),
  ", max=", round(ph_q[[5]], 3)
)
if (mean(pheno_core$mort > 0.9999, na.rm = TRUE) > 0.25) {
  warning("More than 25% of participants have mort > 0.9999 after stabilization. Re-check biomarker units, especially CRP and glucose.")
}

```

# Exposure Processing and LOD Handling

```{r exposure-processing}
pfas_vars <- c("LBXPFOS", "LBXPFOA", "LBXPFNA", "LBXPFHS", "LBXPFDA")

# --- ISOMER HARMONIZATION (2013-2018) ---
# NHANES split PFOS and PFOA into linear and branched isomers.
# We must apply the LOD/sqrt(2) rule to the individual isomers FIRST, then sum them.

isomer_map <- list(
  pfos_lin = c("SSNPFOS", "LBXNFOS"),
  pfos_br  = c("SSMPFOS", "LBXMFOS"),
  pfoa_lin = c("SSNPFOA", "LBXNFOA"),
  pfoa_br  = c("SSBPFOA", "LBXBFOA")
)

lc_map <- list(
  pfos_lin = c("SDNPFOSL", "LBDNFOSL"),
  pfos_br  = c("SDMPFOSL", "LBDMFOSL"),
  pfoa_lin = c("SDNPFOAL", "LBDNFOAL"),
  pfoa_br  = c("SDBPFOAL", "LBDBFOAL")
)

for (chem in c("pfos", "pfoa")) {
  # Get available variables for the cycle
  lin_vars <- isomer_map[[paste0(chem, "_lin")]]
  br_vars  <- isomer_map[[paste0(chem, "_br")]]
  lin_lc   <- lc_map[[paste0(chem, "_lin")]]
  br_lc    <- lc_map[[paste0(chem, "_br")]]
  
  lin_val <- coalesce_num(dat, lin_vars)
  br_val  <- coalesce_num(dat, br_vars)
  lin_lc_val <- coalesce_num(dat, lin_lc)
  br_lc_val  <- coalesce_num(dat, br_lc)
  
  # Apply LOD rule: if indicator == 1, divide by sqrt(2)
  lin_val <- ifelse(!is.na(lin_lc_val) & lin_lc_val == 1, lin_val / sqrt(2), lin_val)
  br_val  <- ifelse(!is.na(br_lc_val) & br_lc_val == 1, br_val / sqrt(2), br_val)
  
  # Sum them to recreate the total concentration
  summed_val <- lin_val + br_val
  
  # Inject back into the legacy total variables (LBXPFOS or LBXPFOA)
  target_var <- ifelse(chem == "pfos", "LBXPFOS", "LBXPFOA")
  
  if (target_var %in% names(dat)) {
    dat[[target_var]] <- coalesce_num(dat, c(target_var))
    dat[[target_var]] <- ifelse(is.na(dat[[target_var]]) & !is.na(summed_val), summed_val, dat[[target_var]])
  } else {
    dat[[target_var]] <- summed_val
  }
}
# --- END ISOMER HARMONIZATION ---

blood_metal_vars <- c("LBXBPB", "LBXBCD", "LBXTHG")
urine_metal_vars <- c("URXUAS", "URXUCD", "URXUPB")

chosen_exposures <- switch(
  config$exposure_family,
  pfas = pfas_vars,
  metals = c(blood_metal_vars, urine_metal_vars),
  both = c(pfas_vars, blood_metal_vars, urine_metal_vars),
  pfas_vars
)
chosen_exposures <- chosen_exposures[chosen_exposures %in% names(dat)]
if (length(chosen_exposures) < 3) stop("Too few exposure variables found; choose different exposure family/cycles.")

for (v in chosen_exposures) {
  vv <- apply_lod_rule(dat, v)
  dat[[paste0(v, "_log2")]] <- log2(vv)
}

if (any(urine_metal_vars %in% chosen_exposures)) {
  for (uv in intersect(urine_metal_vars, chosen_exposures)) {
    vv <- apply_lod_rule(dat, uv)
    dat[[paste0(uv, "_ucr_adj_log2")]] <- log2((vv / pmax(dat$ucr, 1e-6)) * 100)
  }
}

exposure_vars <- names(dat)[str_detect(names(dat), "_log2$|_ucr_adj_log2$")]
exposure_vars <- unique(exposure_vars)
# Prevent collinear double-counting: use creatinine-adjusted urine metals only, drop raw _log2 for them.
raw_urine_vars <- paste0(urine_metal_vars, "_log2")
exposure_vars <- setdiff(exposure_vars, raw_urine_vars)
# Primary analysis: fixed 4-PFAS subset common across all retained cycles. Sensitivity: available-exposure (adaptive) mixture run as secondary.
primary_exposure_vars <- c("LBXPFOS_log2", "LBXPFOA_log2", "LBXPFNA_log2", "LBXPFHS_log2")
primary_exposure_vars <- primary_exposure_vars[primary_exposure_vars %in% exposure_vars]
adaptive_exposure_vars <- exposure_vars

# LOD diagnostic table -- computed on dat (post-merge, pre-log transform) stratified by cycle.
lod_diag <- lapply(chosen_exposures, function(v) {
  x_raw <- suppressWarnings(as.numeric(as.character(dat[[v]])))
  base <- sub("^(LBX|URX)", "", v)
  lc_candidates <- c(
    paste0(v, "LC"), paste0("LBD", base, "LC"), paste0("URD", base, "LC"),
    paste0(v, "L"),  paste0("LBD", base, "L"),  paste0("URD", base, "L")
  )
  lc_var <- lc_candidates[lc_candidates %in% names(dat)]
  has_lc <- length(lc_var) > 0
  if (has_lc) {
    lc_vals <- suppressWarnings(as.numeric(as.character(dat[[lc_var[1]]])))
    below <- !is.na(lc_vals) & lc_vals == 1
  } else {
    below <- rep(FALSE, length(x_raw))
  }

  by_cycle <- dat %>%
    mutate(.x = x_raw, .below = below) %>%
    group_by(cycle) %>%
    summarise(
      n = sum(!is.na(.x)),
      n_below = sum(.below, na.rm = TRUE),
      pct_below = if (sum(!is.na(.x)) > 0) round(sum(.below, na.rm = TRUE) / sum(!is.na(.x)) * 100, 1) else NA_real_,
      median_below = if (sum(.below & !is.na(.x)) > 0) median(.x[.below], na.rm = TRUE) else NA_real_,
      median_above = if (sum(!.below & !is.na(.x)) > 0) median(.x[!.below & !is.na(.x)], na.rm = TRUE) else NA_real_,
      min_positive = if (any(.x > 0, na.rm = TRUE)) min(.x[.x > 0], na.rm = TRUE) else NA_real_,
      .groups = "drop"
    ) %>%
    mutate(exposure = v, lc_indicator = if (has_lc) lc_var[1] else "none") %>%
    select(exposure, cycle, lc_indicator, everything())
  by_cycle
}) %>% bind_rows()
kable(lod_diag, caption = "LOD diagnostic per exposure by cycle. Verify: median_below ~ LOD value; if it diverges from min_positive, review the LC->x/sqrt(2) assumption.")
```

# Analytic Sample and Survey Design

```{r analytic-sample}
analytic <- dat %>%
  filter(age >= config$min_age)

# PFAS workflow: harmonize rotating 2-year subsample weights into one unified PFAS weight (WTPFAS2YR).
analytic <- analytic %>%
  mutate(
    WTPFAS2YR = coalesce_num(cur_data(), c("WTSA2YR", "WTSB2YR", "WTSC2YR", "WTSAF2YR"))
  )

weight_candidates <- names(analytic)[stringr::str_detect(names(analytic), "^WT")]
kable(data.frame(weight_variable = weight_candidates), caption = "Weight audit: all WT* variables in analytic dataset.")

# Exposure-specific weight audit in the exposure-complete analytic sample (fixed primary PFAS set).
complete_expo <- analytic %>%
  mutate(.has_all_expo = if_all(all_of(primary_exposure_vars), ~ !is.na(.x)))
wt_cols <- names(analytic)[str_detect(names(analytic), "^WT")]
# PFAS auto-selection eligibility: exclude full-sample interview/MEC weights; keep PFAS-appropriate subsample candidates.
if (tolower(config$exposure_family) == "pfas") {
  wt_cols <- setdiff(wt_cols, c("WTINT2YR", "WTMEC2YR"))
}
if (length(wt_cols) > 0) {
  wt_audit <- complete_expo %>%
    filter(.has_all_expo) %>%
    summarise(across(all_of(wt_cols), ~ round(mean(!is.na(.x)) * 100, 1))) %>%
    pivot_longer(everything(), names_to = "weight", values_to = "pct_nonmissing_complete_expo") %>%
    arrange(desc(pct_nonmissing_complete_expo))
  kable(wt_audit, caption = "Weight availability (%) among participants with complete exposure set. Use the weight with highest coverage that matches your subsample.")
}

# Auto-weight: rank eligible candidate weights by coverage and apply the top PFAS-appropriate subsample weight.
if (config$weight_type == "auto" && exists("wt_audit") && nrow(wt_audit) > 0) {
  wt_audit_auto <- wt_audit
  if (tolower(config$exposure_family) == "pfas") {
    wt_audit_auto <- wt_audit_auto %>% filter(!(weight %in% c("WTINT2YR", "WTMEC2YR")))
  }
  if (nrow(wt_audit_auto) == 0) {
    stop("Auto-weight selection failed for PFAS: no eligible WTS* subsample weights were available in the audit table.")
  }
  top_row <- wt_audit_auto %>% slice(1)
  config$weight_type <- top_row$weight
  message(">>> Auto-weight resolved: Applying '", config$weight_type, "' (Coverage: ", top_row$pct_nonmissing_complete_expo, "%)")
}

weight_info <- resolve_weight_var(analytic, config$weight_type, config$exposure_family)

analytic$weight_2yr <- suppressWarnings(as.numeric(as.character(analytic[[weight_info$var]])))

analytic <- analytic %>%
  filter(!is.na(phenoage), !is.na(age), !is.na(ckd), !is.na(weight_2yr), !is.na(strata), !is.na(psu))

n_cycles <- length(unique(analytic$cycle))
# After resolving the PFAS-appropriate 2-year weight, pooled analyses divide by the number of contributing cycles in the final sample (NCHS multi-cycle guidance).
n_cycles_exposure_available <- analytic %>%
  filter(if_any(all_of(primary_exposure_vars), ~ !is.na(.x))) %>%
  summarise(n = n_distinct(cycle)) %>%
  pull(n)
if (!is.finite(n_cycles_exposure_available) || n_cycles_exposure_available < 1) {
  n_cycles_exposure_available <- n_cycles
}
repool_weights <- function(df) {
  ncy <- dplyr::n_distinct(df$cycle[!is.na(df$weight_2yr)])
  if (!is.finite(ncy) || ncy < 1) stop("Invalid cycle count for weight pooling.")
  df$weight_combined <- df$weight_2yr / ncy
  df
}
analytic <- repool_weights(analytic)
message("Weight pooling: repool_weights() applied to analytic (n_cycles in sample = ", n_cycles_exposure_available, ")")

# Audit B: weight availability in the final analytic sample (after all exclusions).
if (length(wt_cols) > 0) {
  wt_audit_final <- analytic %>%
    summarise(across(any_of(wt_cols), ~ round(mean(!is.na(.x)) * 100, 1))) %>%
    pivot_longer(everything(), names_to = "weight", values_to = "pct_nonmissing_final_sample") %>%
    arrange(desc(pct_nonmissing_final_sample))
  kable(wt_audit_final, caption = "Weight availability (%) in final analytic sample (post-exclusion). This is the definitive audit.")
}

# PRIMARY: survey-weighted residualization of PhenoAge on age.
# This ensures PhenoAgeAccel is a population-representative construct.
analytic$phenoage_accel_unweighted <- NA_real_
fit_age_uw <- lm(phenoage ~ age, data = analytic)
analytic$phenoage_accel_unweighted <- residuals(fit_age_uw)
keep_w <- which(!is.na(analytic$phenoage) & !is.na(analytic$age) & !is.na(analytic$weight_combined))
if (length(keep_w) > 0) {
  svy_resid <- svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined, nest = TRUE,
                          data = analytic[keep_w, , drop = FALSE])
  fit_age_wt <- svyglm(phenoage ~ age, design = svy_resid, family = gaussian())
  pred_wt <- as.numeric(predict(fit_age_wt))
  analytic$phenoage_accel <- NA_real_
  analytic$phenoage_accel[keep_w] <- analytic$phenoage[keep_w] - pred_wt
}
analytic <- analytic %>% filter(!is.na(phenoage_accel))
message(">>> PhenoAgeAccel: primary = survey-weighted residualization; unweighted version retained as sensitivity.")

ckd_marker_availability <- tibble::tibble(
  group = c("Both eGFR and ACR", "eGFR only", "ACR only", "Neither"),
  N = c(
    sum(!is.na(analytic$egfr) & !is.na(analytic$acr)),
    sum(!is.na(analytic$egfr) & is.na(analytic$acr)),
    sum(is.na(analytic$egfr) & !is.na(analytic$acr)),
    sum(is.na(analytic$egfr) & is.na(analytic$acr))
  )
)

if (!all(c(0, 1) %in% unique(analytic$ckd))) {
  stop("Both CKD groups are not represented after exclusions.")
}

svy <- svydesign(
  id = ~psu, strata = ~strata, weights = ~weight_combined, nest = TRUE, data = analytic
)

kable(data.frame(weight_selected = weight_info$var, note = weight_info$note), caption = "Selected NHANES weight variable")
kable(ckd_marker_availability, caption = "CKD marker availability in analytic sample")
```

# Mixture Modeling

```{r mixture-models}
valid_num <- function(v, d) v %in% names(d) && length(unique(stats::na.omit(d[[v]]))) >= 2
valid_cat <- function(v, d) v %in% names(d) && length(unique(stats::na.omit(as.character(d[[v]])))) >= 2

# Primary adjustment: lean set to avoid overadjustment/mediator bias (diabetes, HTN, CVD may be on pathway).
# Aligned with EHP/Env Int mixture papers: sex, race, PIR, BMI, smoking, cycle; plus alcohol for confounding control.
# Smoking and alcohol are 3-level factors (Never/Former/Current) for consistency with Table 1.
clinical_covars <- c("diabetes", "hypertension", "cvd")  # used only in sensitivity "cardiometabolic"
# No age: outcome is phenoage_accel (age-residualized). Same covariates for qgcomp and Model B.
covars <- c(
  if (valid_cat("sex", analytic)) "factor(sex)",
  if (valid_cat("race_eth", analytic)) "factor(race_eth)",
  if (valid_num("pir", analytic)) "pir",
  if (valid_cat("smoker3", analytic)) "factor(smoker3)",
  if (valid_num("bmi", analytic)) "bmi",
  if (valid_cat("alcohol3", analytic)) "factor(alcohol3)",
  if (valid_cat("cycle", analytic)) "factor(cycle)"
)

# Sanity checks before modeling.
message("--- Pre-modeling sanity checks ---")
message("Analytic N: ", nrow(analytic))
message("PhenoAgeAccel: mean=", round(mean(analytic$phenoage_accel, na.rm = TRUE), 3),
        ", SD=", round(sd(analytic$phenoage_accel, na.rm = TRUE), 3))
message("CKD prevalence: ", round(mean(analytic$ckd == 1, na.rm = TRUE) * 100, 1), "%")
message("Weight selected: ", weight_info$var)

# qgcomp complete-case: use primary (fixed) exposure vars with sufficient variability and non-missing in >= 50% of analytic.
qg_min_frac <- 0.5
qg_exposure_vars_for_cc <- primary_exposure_vars[vapply(primary_exposure_vars, function(v) {
  x <- suppressWarnings(as.numeric(analytic[[v]]))
  finite <- is.finite(x)
  n_ok <- sum(finite)
  pct_ok <- n_ok / max(1, nrow(analytic))
  n_ok > 0 && dplyr::n_distinct(x[finite]) >= 2 && pct_ok >= qg_min_frac
}, logical(1))]
if (length(qg_exposure_vars_for_cc) < 2) {
  qg_exposure_vars_for_cc <- primary_exposure_vars[vapply(primary_exposure_vars, function(v) {
    x <- suppressWarnings(as.numeric(analytic[[v]]))
    x <- x[is.finite(x)]
    length(x) > 0 && dplyr::n_distinct(x) >= 2
  }, logical(1))]
  if (length(qg_exposure_vars_for_cc) < 2) qg_exposure_vars_for_cc <- primary_exposure_vars
  message("qgcomp: few primary exposures met variance + ", round(qg_min_frac * 100), "% non-missing; using variance-only subset or all primary_exposure_vars.")
}
message("qgcomp exposures for complete-case (primary fixed set, variance + >= ", round(qg_min_frac * 100), "% non-missing): ", paste(qg_exposure_vars_for_cc, collapse = ", "))
# Qgcomp uses same adjustment as Model B (sex, race, PIR, BMI, smoking [3-level], alcohol [3-level], cycle); no age (outcome is age-residualized).
rhs_qg_cc <- paste(c(qg_exposure_vars_for_cc, covars), collapse = " + ")
f_qg_cc <- as.formula(paste("phenoage_accel ~", rhs_qg_cc))
qg_cc <- complete.cases(model.frame(f_qg_cc, data = analytic, na.action = na.pass))
analytic_qg <- analytic[qg_cc, , drop = FALSE]
analytic_qg <- repool_weights(analytic_qg)
# Normalize precision weights for qgcomp so glm does not treat sum(weights) as N (avoids N = 300M, infinitesimal SE).
analytic_qg$weight_normalized <- analytic_qg$weight_combined / mean(analytic_qg$weight_combined, na.rm = TRUE)
# Ensure covariates are proper type for qgcomp (character -> wrong model matrix, psi ~ 0).
for (v in c("phenoage_accel", "age", "pir", "bmi")) {
  if (v %in% names(analytic_qg)) analytic_qg[[v]] <- suppressWarnings(as.numeric(as.character(analytic_qg[[v]])))
}
analytic_qg$sex      <- factor(as.character(analytic_qg$sex))
analytic_qg$race_eth <- factor(as.character(analytic_qg$race_eth))
analytic_qg$cycle    <- factor(as.character(analytic_qg$cycle))
if ("smoker3"  %in% names(analytic_qg)) analytic_qg$smoker3  <- factor(as.character(analytic_qg$smoker3),  levels = c("Never", "Former", "Current"))
if ("alcohol3" %in% names(analytic_qg)) analytic_qg$alcohol3 <- factor(as.character(analytic_qg$alcohol3), levels = c("Never", "Former", "Current"))
for (v in qg_exposure_vars_for_cc) {
  if (v %in% names(analytic_qg)) analytic_qg[[v]] <- suppressWarnings(as.numeric(as.character(analytic_qg[[v]])))
}
message("qgcomp complete-case: same covariates as Model B; N = ", nrow(analytic_qg))

# Among those, keep only exposures that still have variance in the complete-case sample (in case one collapses).
qg_exposure_vars <- qg_exposure_vars_for_cc[vapply(qg_exposure_vars_for_cc, function(v) {
  x <- suppressWarnings(as.numeric(analytic_qg[[v]]))
  x <- x[is.finite(x)]
  length(x) > 0 && dplyr::n_distinct(x) >= 2
}, logical(1))]
if (length(qg_exposure_vars) < 2) {
  warning("Too few finite/variable exposures for qgcomp after complete-case filtering; Table 2 will be NA.")
}
if (length(qg_exposure_vars) > 0) {
  analytic_qg <- analytic_qg %>%
    filter(if_all(all_of(c("phenoage_accel", qg_exposure_vars)), ~ is.finite(.x)))
}
q_qg <- as.integer(config$q)
if (!is.finite(q_qg) || q_qg < 2) stop("config$q must be an integer >= 2.")

# Qgcomp model: same adjustment as Model B (no age).
covars_qg <- c(
  if (valid_cat("sex", analytic_qg)) "factor(sex)",
  if (valid_cat("race_eth", analytic_qg)) "factor(race_eth)",
  if (valid_num("pir", analytic_qg)) "pir",
  if (valid_cat("smoker3", analytic_qg)) "factor(smoker3)",
  if (valid_num("bmi", analytic_qg)) "bmi",
  if (valid_cat("alcohol3", analytic_qg)) "factor(alcohol3)",
  if (valid_cat("cycle", analytic_qg)) "factor(cycle)"
)
qg_covars_intended <- covars_qg
build_qg_formula <- function(exps, covs) {
  rhs <- paste(c(exps, covs), collapse = " + ")
  as.formula(paste("phenoage_accel ~", rhs))
}
f_qg <- build_qg_formula(qg_exposure_vars, covars_qg)

repeat {
  mf_qg <- tryCatch(model.frame(f_qg, data = analytic_qg, na.action = na.omit), error = function(e) NULL)
  if (is.null(mf_qg) || nrow(mf_qg) == 0) break
  factor_terms_qg <- names(mf_qg)[vapply(mf_qg, is.factor, logical(1))]
  bad_factor_terms_qg <- factor_terms_qg[vapply(mf_qg[factor_terms_qg], function(x) nlevels(droplevels(x)) < 2, logical(1))]
  removable_qg <- intersect(covars_qg, bad_factor_terms_qg)
  if (length(removable_qg) == 0) break
  message("qgcomp overall: dropping collapsed factor term(s): ", paste(removable_qg, collapse = ", "))
  covars_qg <- setdiff(covars_qg, removable_qg)
  f_qg <- build_qg_formula(qg_exposure_vars, covars_qg)
}
qg_covars_realized <- covars_qg
message("qgcomp exposures used: ", paste(qg_exposure_vars, collapse = ", "))
message("qgcomp finite analytic N: ", nrow(analytic_qg))

qgcomp_overall_error <- NULL
fit_qg_overall <- tryCatch(
  qgcomp::qgcomp.noboot(
    f = f_qg, expnms = qg_exposure_vars, data = analytic_qg,
    family = gaussian(), q = q_qg, weights = analytic_qg$weight_normalized
  ),
  error = function(e) {
    qgcomp_overall_error <<- conditionMessage(e)
    NULL
  }
)
if (!is.null(fit_qg_overall)) {
  message("qgcomp N used: ", fit_qg_overall$fit$n[1])
} else {
  warning("qgcomp overall model failed: ", qgcomp_overall_error)
}

# Survey-weighted interaction model using quantized summed score.
# Quantile cutpoints are computed within each NHANES cycle using cycle-specific sampling weights.
qdat <- analytic
qdat$ckd <- as.integer(qdat$ckd)
# Strict sample = complete-case on qgcomp-compatible PFAS subset (same as analytic_qg). Label reflects this.
qdat_strict <- analytic_qg
qdat_strict$ckd <- as.integer(qdat_strict$ckd)

# Primary model: fixed PFAS set (primary_exposure_vars). Restrict to complete case on primary set.
qdat <- qdat %>% filter(if_all(all_of(primary_exposure_vars), ~ !is.na(.x)))
if (config$require_complete_exposure_set) {
  # Optional extra: could also require full adaptive set here; we use primary only for main.
}
message("Primary interaction: fixed PFAS set = ", paste(primary_exposure_vars, collapse = ", "))
message("N primary pre-quantile sample: ", nrow(qdat))
message("N strict sample (Strict complete-case, qgcomp-compatible PFAS subset): ", nrow(qdat_strict))

weighted_quantile <- function(x, w, probs) {
  keep <- is.finite(x) & is.finite(w) & w > 0
  x <- x[keep]; w <- w[keep]
  if (length(x) == 0) return(rep(NA_real_, length(probs)))
  ord <- order(x)
  x <- x[ord]; w <- w[ord]
  cw <- cumsum(w) / sum(w)
  sapply(probs, function(p) {
    idx <- which(cw >= p)[1]
    if (is.na(idx)) x[length(x)] else x[idx]
  })
}

# Primary mixture score: fixed set, cycle-specific quantiles to avoid temporal confounding (PFAS trends across retained cycles).
for (v in primary_exposure_vars) {
  qdat[[paste0(v, "_q")]] <- NA_real_
  for (cy in unique(qdat$cycle)) {
    idx <- qdat$cycle == cy
    x_cy <- suppressWarnings(as.numeric(qdat[[v]][idx]))
    w_cy <- suppressWarnings(as.numeric(qdat$weight_combined[idx]))
    probs <- weighted_quantile(x_cy, w_cy, probs = seq(0, 1, length.out = config$q + 1))
    br <- sort(unique(as.numeric(probs)))
    if (length(br) > 1) {
      br[1] <- -Inf
      br[length(br)] <- Inf
    }
    if (length(br) < 3 || any(diff(br) <= 0)) {
      r <- rank(qdat[[v]][idx], na.last = "keep", ties.method = "average")
      qdat[[paste0(v, "_q")]][idx] <- as.numeric(cut(r, breaks = config$q, labels = FALSE, include.lowest = TRUE))
    } else {
      qdat[[paste0(v, "_q")]][idx] <- as.numeric(cut(qdat[[v]][idx], breaks = br, include.lowest = TRUE, labels = FALSE))
    }
  }
}
qvars <- paste0(primary_exposure_vars, "_q")
qdat$nonmissing_q_exposures <- rowSums(!is.na(qdat[, qvars, drop = FALSE]))
qdat$mixture_qscore <- rowMeans(qdat[, qvars, drop = FALSE], na.rm = TRUE)
# Primary: require all primary PFAS non-missing (fixed set).
required_primary <- length(primary_exposure_vars)
qdat$mixture_qscore <- ifelse(qdat$nonmissing_q_exposures < required_primary, NA_real_, qdat$mixture_qscore)
# By-cycle table for reporting (primary set is fixed; this documents it).
cycle_capacity <- tibble(
  cycle = unique(qdat$cycle),
  measured_exposures = length(primary_exposure_vars),
  PFAS_measured = paste(primary_exposure_vars, collapse = ", "),
  required_nonmissing_exposures = required_primary
)

# Build strict mixture score: same exposure set as qgcomp, cycle-specific quantiles.
for (v in qg_exposure_vars) {
  qdat_strict[[paste0(v, "_q")]] <- NA_real_
  for (cy in unique(qdat_strict$cycle)) {
    idx <- qdat_strict$cycle == cy
    x_cy <- suppressWarnings(as.numeric(qdat_strict[[v]][idx]))
    w_cy <- suppressWarnings(as.numeric(qdat_strict$weight_combined[idx]))
    probs_s <- weighted_quantile(x_cy, w_cy, probs = seq(0, 1, length.out = config$q + 1))
    br_s <- sort(unique(as.numeric(probs_s)))
    if (length(br_s) > 1) {
      br_s[1] <- -Inf
      br_s[length(br_s)] <- Inf
    }
    if (length(br_s) < 3 || any(diff(br_s) <= 0)) {
      r_s <- rank(qdat_strict[[v]][idx], na.last = "keep", ties.method = "average")
      qdat_strict[[paste0(v, "_q")]][idx] <- as.numeric(cut(r_s, breaks = config$q, labels = FALSE, include.lowest = TRUE))
    } else {
      qdat_strict[[paste0(v, "_q")]][idx] <- as.numeric(cut(qdat_strict[[v]][idx], breaks = br_s, include.lowest = TRUE, labels = FALSE))
    }
  }
}
qvars_strict <- paste0(qg_exposure_vars, "_q")
qdat_strict$nonmissing_q_exposures <- rowSums(!is.na(qdat_strict[, qvars_strict, drop = FALSE]))
qdat_strict$mixture_qscore <- rowMeans(qdat_strict[, qvars_strict, drop = FALSE], na.rm = TRUE)

# Hard type enforcement to prevent silent coercion issues in svyglm.
qdat <- qdat %>%
  mutate(
    phenoage_accel = suppressWarnings(as.numeric(phenoage_accel)),
    mixture_qscore = suppressWarnings(as.numeric(mixture_qscore)),
    weight_combined = suppressWarnings(as.numeric(weight_combined)),
    psu = suppressWarnings(as.numeric(psu)),
    strata = suppressWarnings(as.numeric(strata)),
    ckd = as.integer(ckd)
  )
qdat_strict <- qdat_strict %>%
  mutate(
    phenoage_accel = suppressWarnings(as.numeric(phenoage_accel)),
    mixture_qscore = suppressWarnings(as.numeric(mixture_qscore)),
    weight_combined = suppressWarnings(as.numeric(weight_combined)),
    psu = suppressWarnings(as.numeric(psu)),
    strata = suppressWarnings(as.numeric(strata)),
    ckd = as.integer(ckd)
  )

# Primary model: fixed PFAS set (same in every cycle).
cycle_capacity_with_min <- cycle_capacity
kable(
  cycle_capacity_with_min,
  caption = "Primary analysis: fixed PFAS set by cycle. Same chemicals and required non-missing in all cycles."
)

kable(
  tibble(
    statistic = c("N", "Mean", "SD", "Min", "P25", "Median", "P75", "Max"),
    value = c(
      sum(!is.na(qdat$mixture_qscore)),
      round(mean(qdat$mixture_qscore, na.rm = TRUE), 3),
      round(sd(qdat$mixture_qscore, na.rm = TRUE), 3),
      round(min(qdat$mixture_qscore, na.rm = TRUE), 3),
      round(quantile(qdat$mixture_qscore, 0.25, na.rm = TRUE), 3),
      round(median(qdat$mixture_qscore, na.rm = TRUE), 3),
      round(quantile(qdat$mixture_qscore, 0.75, na.rm = TRUE), 3),
      round(max(qdat$mixture_qscore, na.rm = TRUE), 3)
    )
  ),
  caption = paste0("Primary mixture quantile score (fixed set, N chemicals = ", length(primary_exposure_vars),
                   "). Score = mean quantile bin across ", paste(primary_exposure_vars, collapse = ", "),
                   ". Complete case on primary set required.")
)

stopifnot(all(c("psu", "strata", "weight_combined") %in% names(qdat)))
qdat <- qdat %>% filter(!is.na(psu), !is.na(strata), !is.na(weight_combined), !is.na(phenoage_accel), !is.na(ckd), !is.na(mixture_qscore))
qdat_strict <- qdat_strict %>% filter(!is.na(psu), !is.na(strata), !is.na(weight_combined), !is.na(phenoage_accel), !is.na(ckd), !is.na(mixture_qscore))
qdat         <- repool_weights(qdat)
qdat_strict  <- repool_weights(qdat_strict)

# Ensure all model variables are proper type so svyglm does not zero out coefficients (character -> wrong design matrix).
for (v in exposure_vars) {
  if (v %in% names(qdat)) qdat[[v]] <- suppressWarnings(as.numeric(as.character(qdat[[v]])))
}
qdat$phenoage_accel <- suppressWarnings(as.numeric(as.character(qdat$phenoage_accel)))
qdat$age            <- suppressWarnings(as.numeric(as.character(qdat$age)))
qdat$pir            <- suppressWarnings(as.numeric(as.character(qdat$pir)))
qdat$bmi            <- suppressWarnings(as.numeric(as.character(qdat$bmi)))
qdat$sex            <- factor(as.character(qdat$sex))
qdat$race_eth       <- factor(as.character(qdat$race_eth))
qdat$cycle          <- factor(as.character(qdat$cycle))
if ("smoker3"  %in% names(qdat)) qdat$smoker3  <- factor(as.character(qdat$smoker3),  levels = c("Never", "Former", "Current"))
if ("alcohol3" %in% names(qdat)) qdat$alcohol3 <- factor(as.character(qdat$alcohol3), levels = c("Never", "Former", "Current"))
qdat$ckd            <- as.integer(qdat$ckd)

for (v in exposure_vars) {
  if (v %in% names(qdat_strict)) qdat_strict[[v]] <- suppressWarnings(as.numeric(as.character(qdat_strict[[v]])))
}
qdat_strict$phenoage_accel <- suppressWarnings(as.numeric(as.character(qdat_strict$phenoage_accel)))
qdat_strict$age            <- suppressWarnings(as.numeric(as.character(qdat_strict$age)))
qdat_strict$pir            <- suppressWarnings(as.numeric(as.character(qdat_strict$pir)))
qdat_strict$bmi            <- suppressWarnings(as.numeric(as.character(qdat_strict$bmi)))
qdat_strict$sex            <- factor(as.character(qdat_strict$sex))
qdat_strict$race_eth       <- factor(as.character(qdat_strict$race_eth))
qdat_strict$cycle          <- factor(as.character(qdat_strict$cycle))
if ("smoker3"  %in% names(qdat_strict)) qdat_strict$smoker3  <- factor(as.character(qdat_strict$smoker3),  levels = c("Never", "Former", "Current"))
if ("alcohol3" %in% names(qdat_strict)) qdat_strict$alcohol3 <- factor(as.character(qdat_strict$alcohol3), levels = c("Never", "Former", "Current"))
qdat_strict$ckd            <- as.integer(qdat_strict$ckd)

message("---- Model B diagnostics ----")
message("N qdat: ", nrow(qdat))
print(table(qdat$ckd, useNA = "ifany"))
message("N qdat_strict: ", nrow(qdat_strict))
print(table(qdat_strict$ckd, useNA = "ifany"))
sd_all <- sd(qdat$mixture_qscore, na.rm = TRUE)
sd_by <- tapply(qdat$mixture_qscore, qdat$ckd, sd, na.rm = TRUE)
message("SD(mixture_qscore) overall: ", sd_all)
print(sd_by)

if (length(unique(stats::na.omit(qdat$ckd))) < 2) {
  stop("After exposure/design filtering, only one CKD group remains in qdat. ",
       "Interaction mixture_qscore*ckd cannot be estimated. ",
       "Set config$require_complete_exposure_set = FALSE or reduce exposure set/cycles.")
}
if (any(is.finite(sd_by) & sd_by < 1e-8)) {
  stop("mixture_qscore has near-zero variance within at least one CKD group. ",
       "Interaction cannot be estimated; check quantile binning/ties and exposure variability.")
}

miss_compare_ckd <- qdat %>%
  mutate(
    retained_primary = complete.cases(phenoage_accel, mixture_qscore, ckd, age, sex, race_eth, pir, smoker3, bmi, alcohol3, cycle)
  ) %>%
  group_by(ckd, retained_primary) %>%
  summarise(
    N = n(),
    mean_age = round(mean(age, na.rm = TRUE), 1),
    female_pct = round(mean(sex == "Female", na.rm = TRUE) * 100, 1),
    .groups = "drop"
  ) %>%
  group_by(ckd) %>%
  mutate(pct_within_ckd = round(100 * N / sum(N), 1)) %>%
  ungroup() %>%
  mutate(
    ckd = ifelse(ckd == 1, "CKD", "No CKD"),
    retained_primary = ifelse(retained_primary, "Retained in Model B", "Dropped from Model B")
  )
kable(miss_compare_ckd, caption = "Primary interaction model retention by CKD status (retained vs dropped participants).")

# Use explicit data-frame copy so survey design uses correct types (avoids ~0 coefficients from reference/lazy issues).
qdat_svy <- as.data.frame(qdat)
qdat_svy$phenoage_accel <- as.numeric(qdat_svy$phenoage_accel)
qdat_svy$mixture_qscore <- as.numeric(qdat_svy$mixture_qscore)
qdat_svy$ckd            <- as.integer(qdat_svy$ckd)
qdat_svy$age            <- as.numeric(qdat_svy$age)
qdat_svy$pir            <- as.numeric(qdat_svy$pir)
qdat_svy$bmi            <- as.numeric(qdat_svy$bmi)
qdat_svy$sex            <- factor(qdat_svy$sex)
qdat_svy$race_eth       <- factor(qdat_svy$race_eth)
qdat_svy$cycle          <- factor(qdat_svy$cycle)
if ("smoker3"  %in% names(qdat_svy)) qdat_svy$smoker3  <- factor(as.character(qdat_svy$smoker3),  levels = c("Never", "Former", "Current"))
if ("alcohol3" %in% names(qdat_svy)) qdat_svy$alcohol3 <- factor(as.character(qdat_svy$alcohol3), levels = c("Never", "Former", "Current"))
qdat_svy$weight_combined <- as.numeric(qdat_svy$weight_combined)
qdat_svy$psu            <- as.numeric(qdat_svy$psu)
qdat_svy$strata         <- as.numeric(qdat_svy$strata)
# Drop unused factor levels to avoid singularity with full survey design.
qdat_svy$sex      <- droplevels(factor(qdat_svy$sex))
qdat_svy$race_eth <- droplevels(factor(qdat_svy$race_eth))
qdat_svy$cycle    <- droplevels(factor(qdat_svy$cycle))
if ("smoker3"  %in% names(qdat_svy)) qdat_svy$smoker3  <- droplevels(qdat_svy$smoker3)
if ("alcohol3" %in% names(qdat_svy)) qdat_svy$alcohol3 <- droplevels(qdat_svy$alcohol3)

# ---- Pre-compute sensitivity-derived columns on qdat_svy so they ride into MI imputations ----
# These outcomes/redefinitions depend only on biomarkers + age (not imputed), so they can be computed once.
ph2_comp_qsvy <- compute_phenoage_components(qdat_svy, include_scr = FALSE, include_alb = TRUE)
ph2_qsvy <- ph2_comp_qsvy$phenoage
qdat_svy$phenoage_nokid <- NA_real_
keep2_qsvy <- which(!is.na(ph2_qsvy) & !is.na(qdat_svy$age))
if (length(keep2_qsvy) > 0) {
  fit_nokid_qsvy <- lm(ph2_qsvy[keep2_qsvy] ~ qdat_svy$age[keep2_qsvy])
  qdat_svy$phenoage_nokid[keep2_qsvy] <- residuals(fit_nokid_qsvy)
}
ph3_comp_qsvy <- compute_phenoage_components(qdat_svy, include_scr = FALSE, include_alb = FALSE)
ph3_qsvy <- ph3_comp_qsvy$phenoage
qdat_svy$phenoage_nokid_both <- NA_real_
keep3_qsvy <- which(!is.na(ph3_qsvy) & !is.na(qdat_svy$age))
if (length(keep3_qsvy) > 0) {
  fit_nokid_both_qsvy <- lm(ph3_qsvy[keep3_qsvy] ~ qdat_svy$age[keep3_qsvy])
  qdat_svy$phenoage_nokid_both[keep3_qsvy] <- residuals(fit_nokid_both_qsvy)
}
qdat_svy$ckd_sens <- ifelse(!is.na(qdat_svy$egfr), as.integer(qdat_svy$egfr < 60), NA_integer_)

# Mirror the same derivations on qdat for complete-case sensitivities.
qdat$phenoage_nokid       <- qdat_svy$phenoage_nokid[match(qdat$SEQN, qdat_svy$SEQN)]
qdat$phenoage_nokid_both  <- qdat_svy$phenoage_nokid_both[match(qdat$SEQN, qdat_svy$SEQN)]
qdat$ckd_sens             <- qdat_svy$ckd_sens[match(qdat$SEQN, qdat_svy$SEQN)]

# Full complex survey design (PSU/strata) for valid inference; droplevels above avoids empty cells.
svy_q <- tryCatch(
  svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined, nest = TRUE, data = qdat_svy),
  error = function(e) {
    message("Full design failed (", conditionMessage(e), "); falling back to weights-only design.")
    svydesign(id = ~1, strata = NULL, weights = ~weight_combined, data = qdat_svy)
  }
)

# ---- Multiple imputation (MICE) for missing model covariates ----
# We NEVER impute the outcome (phenoage_accel), exposures (PFAS / mixture_qscore), CKD, or design vars.
# Imputation base = qdat_svy (already requires non-missing PhenoAge, CKD, weights/PSU/strata, and full PFAS for mixture_qscore).
mice_target_vars <- intersect(
  c("smoker3", "alcohol3", "bmi", "pir", "education",
    "phys_active", "diabetes", "hypertension", "cvd"),
  names(qdat_svy)
)
mice_predictor_vars <- intersect(
  c("phenoage_accel", "ckd", "mixture_qscore", "age", "sex", "race_eth", "cycle"),
  names(qdat_svy)
)

mids_obj <- NULL
imp_list <- NULL
svy_q_mi <- NULL
mice_pre_n_missing <- if (length(mice_target_vars) > 0) {
  vapply(mice_target_vars, function(v) sum(is.na(qdat_svy[[v]])), integer(1))
} else integer(0)

run_mice <- isTRUE(config$use_mice) && length(mice_target_vars) > 0 && any(mice_pre_n_missing > 0)
if (run_mice) {
  mi_data <- qdat_svy[, unique(c(mice_target_vars, mice_predictor_vars)), drop = FALSE]
  # Force factor types correctly for mice methods.
  for (v in c("smoker3", "alcohol3")) {
    if (v %in% names(mi_data)) mi_data[[v]] <- factor(as.character(mi_data[[v]]),
                                                       levels = c("Never", "Former", "Current"))
  }
  for (v in intersect(c("sex", "race_eth", "cycle", "education"), names(mi_data))) {
    mi_data[[v]] <- factor(as.character(mi_data[[v]]))
  }
  for (v in intersect(c("phys_active", "diabetes", "hypertension", "cvd"), names(mi_data))) {
    mi_data[[v]] <- factor(as.character(mi_data[[v]]))
  }
  for (v in intersect(c("phenoage_accel", "mixture_qscore", "age", "bmi", "pir"), names(mi_data))) {
    mi_data[[v]] <- as.numeric(mi_data[[v]])
  }
  if ("ckd" %in% names(mi_data)) mi_data$ckd <- factor(as.character(mi_data$ckd))

  pred_mat <- matrix(0, nrow = ncol(mi_data), ncol = ncol(mi_data),
                     dimnames = list(names(mi_data), names(mi_data)))
  for (v in mice_target_vars) {
    others <- setdiff(names(mi_data), v)
    pred_mat[v, others] <- 1L
  }

  meth <- character(ncol(mi_data)); names(meth) <- names(mi_data)
  for (v in names(mi_data)) {
    if (!(v %in% mice_target_vars)) { meth[v] <- ""; next }
    if (is.factor(mi_data[[v]])) {
      meth[v] <- if (nlevels(droplevels(mi_data[[v]])) == 2) "logreg" else "polyreg"
    } else {
      meth[v] <- "pmm"
    }
  }
  message(">>> MICE: imputing ", paste(mice_target_vars, collapse = ", "),
          " with m=", config$mice_m, ", maxit=", config$mice_maxit)
  mids_obj <- tryCatch(
    mice::mice(mi_data, m = config$mice_m, maxit = config$mice_maxit,
               method = meth, predictorMatrix = pred_mat,
               printFlag = FALSE, seed = config$mice_seed),
    error = function(e) { message("mice() failed: ", conditionMessage(e)); NULL }
  )
  if (!is.null(mids_obj)) {
    imp_list <- mitools::imputationList(lapply(seq_len(config$mice_m), function(i) {
      df <- qdat_svy
      completed <- mice::complete(mids_obj, i)
      for (v in mice_target_vars) df[[v]] <- completed[[v]]
      # Re-pin factor levels in the merged frame so model.frame is consistent across imputations.
      if ("smoker3"  %in% names(df)) df$smoker3  <- factor(as.character(df$smoker3),  levels = c("Never", "Former", "Current"))
      if ("alcohol3" %in% names(df)) df$alcohol3 <- factor(as.character(df$alcohol3), levels = c("Never", "Former", "Current"))
      df
    }))
    svy_q_mi <- tryCatch(
      survey::svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined,
                        nest = TRUE, data = imp_list),
      error = function(e) survey::svydesign(id = ~1, strata = NULL,
                                            weights = ~weight_combined, data = imp_list)
    )
  }
}
n_mi_base <- nrow(qdat_svy)
mice_diag <- tibble::tibble(
  variable = mice_target_vars,
  n_missing_pre = as.integer(mice_pre_n_missing),
  pct_missing_pre = round(100 * mice_pre_n_missing / max(1, n_mi_base), 2)
)
if (length(mice_target_vars) > 0) {
  kable(mice_diag, caption = paste0(
    "MICE pre-imputation missingness (within Model B base sample, N = ",
    format(n_mi_base, big.mark = ","), ")."
  ))
}
mice_active <- isTRUE(config$use_mice) && !is.null(svy_q_mi)
message(">>> MICE active for inference: ", mice_active,
        if (mice_active) paste0(" (m=", config$mice_m, ", N=", n_mi_base, ")") else "")

# ---- Helpers: tidy + svycontrast pooling that handle single-fit OR MI list ----
fit_svyglm_pool <- function(formula, design, family = stats::gaussian()) {
  if (inherits(design, "svyimputationList") || inherits(design, "svy.designs.imputationList")) {
    fits <- with(design, svyglm(formula, family = family))
    pooled <- mitools::MIcombine(fits)
    list(fits = fits, pooled = pooled, mi = TRUE)
  } else {
    fit <- svyglm(formula, design = design, family = family)
    list(fit = fit, mi = FALSE)
  }
}

tidy_pool <- function(x, conf.level = 0.95) {
  alpha <- 1 - conf.level
  if (isTRUE(x$mi)) {
    p <- x$pooled
    coefs <- coef(p); ses <- sqrt(diag(vcov(p)))
    df_v <- if (!is.null(p$df) && length(p$df) == length(coefs)) as.numeric(p$df) else rep(Inf, length(coefs))
    df_v[!is.finite(df_v) | df_v <= 0] <- Inf
    t_val <- coefs / ses
    p_val <- 2 * pt(-abs(t_val), df = df_v)
    crit <- qt(1 - alpha / 2, df = df_v)
    tibble::tibble(
      term = names(coefs),
      estimate = unname(coefs),
      std.error = unname(ses),
      statistic = unname(t_val),
      p.value = unname(p_val),
      conf.low = unname(coefs - crit * ses),
      conf.high = unname(coefs + crit * ses)
    )
  } else {
    broom::tidy(x$fit, conf.int = TRUE, conf.level = conf.level)
  }
}

coef_pool <- function(x) {
  if (isTRUE(x$mi)) coef(x$pooled) else coef(x$fit)
}

contrast_pool <- function(x, c_vec, conf.level = 0.95) {
  alpha <- 1 - conf.level
  if (isTRUE(x$mi)) {
    contrasts <- lapply(x$fits, function(f) survey::svycontrast(f, c_vec))
    pooled <- mitools::MIcombine(contrasts)
    est <- as.numeric(coef(pooled)); se <- as.numeric(sqrt(diag(vcov(pooled))))
    df_v <- if (!is.null(pooled$df) && length(pooled$df) == length(est)) as.numeric(pooled$df) else Inf
    df_v[!is.finite(df_v) | df_v <= 0] <- Inf
  } else {
    ctr <- survey::svycontrast(x$fit, c_vec)
    est <- as.numeric(coef(ctr)); se <- as.numeric(SE(ctr))
    df_v <- x$fit$df.residual
    if (is.null(df_v) || !is.finite(df_v)) df_v <- Inf
  }
  t_val <- est / se
  list(
    estimate = est, se = se, df = df_v,
    p.value = 2 * pt(-abs(t_val), df = df_v),
    conf.low = est - qt(1 - alpha / 2, df = df_v) * se,
    conf.high = est + qt(1 - alpha / 2, df = df_v) * se
  )
}

# ---- qgcomp Rubin pooling: run qgcomp.noboot on each imputation, pool psi (overall mixture coef) ----
# Always writes a small diagnostic log next to the Rmd so failures are visible even when
# knitr suppresses message()/warning() output.
qgcomp_mi_pool <- function(formula, expnms, imp_list_obj, q, weight_col = "weight_combined",
                           log_path = "qgcomp_mi_pool_log.txt") {
  log_lines <- character(0)
  log <- function(...) log_lines <<- c(log_lines, paste0(...))
  on.exit({
    try(writeLines(log_lines, log_path), silent = TRUE)
  }, add = TRUE)
  log("[", format(Sys.time()), "] qgcomp_mi_pool called.")
  log("  formula: ", deparse(formula))
  log("  expnms : ", paste(expnms, collapse = ", "))
  log("  q      : ", q)
  log("  weight : ", weight_col)
  if (is.null(imp_list_obj)) { log("ABORT: imp_list_obj is NULL."); return(NULL) }
  imps <- imp_list_obj$imputations
  m <- length(imps)
  log("  m (#imputations): ", m)
  if (m == 0) { log("ABORT: empty imputation list."); return(NULL) }
  log("  imp[1] cols (first 12): ", paste(head(names(imps[[1]]), 12), collapse = ", "))
  results <- lapply(seq_len(m), function(i) {
    df <- as.data.frame(imps[[i]])
    if (!(weight_col %in% names(df))) {
      log("  imp ", i, ": missing weight column '", weight_col, "'.")
      return(NULL)
    }
    df$weight_normalized <- df[[weight_col]] / mean(df[[weight_col]], na.rm = TRUE)
    needed <- intersect(c("phenoage_accel", expnms, weight_col), names(df))
    df <- df[complete.cases(df[, needed, drop = FALSE]), , drop = FALSE]
    n_unwt <- nrow(df)
    if (n_unwt == 0) {
      log("  imp ", i, ": 0 complete-case rows after filtering on outcome/exposure/weight.")
      return(NULL)
    }
    # Bind data into a temporary global so qgcomp's match.call()-based weights handling
    # resolves the weights vector cleanly outside the lapply closure.
    assign("qgcomp_data_local", df, envir = .GlobalEnv)
    on.exit(if (exists("qgcomp_data_local", envir = .GlobalEnv))
      rm("qgcomp_data_local", envir = .GlobalEnv), add = TRUE)
    fit <- tryCatch(
      do.call(
        qgcomp::qgcomp.noboot,
        list(
          f = formula, expnms = expnms, data = qgcomp_data_local,
          family = stats::gaussian(), q = q,
          weights = qgcomp_data_local$weight_normalized
        )
      ),
      error = function(e) {
        log("  imp ", i, ": qgcomp error: ", conditionMessage(e))
        NULL
      }
    )
    if (is.null(fit)) return(NULL)
    log("  imp ", i, ": OK (n_unwt=", n_unwt,
        ", psi=", tryCatch(round(fit$psi[1], 4), error = function(e) NA), ")")
    list(fit = fit, n_unwt = n_unwt)
  })
  results <- Filter(Negate(is.null), results)
  if (length(results) == 0) {
    log("ABORT: all per-imputation fits failed.")
    return(NULL)
  }
  fits <- lapply(results, `[[`, "fit")
  m_eff <- length(fits)
  psis <- vapply(fits, function(f) unname(f$psi[1]), numeric(1))
  vars <- vapply(fits, function(f) unname(f$var.psi[1]), numeric(1))
  ns   <- vapply(results, function(r) as.integer(r$n_unwt), integer(1))
  qbar <- mean(psis)
  ubar <- mean(vars)
  Bvar <- if (m_eff > 1) stats::var(psis) else 0
  Tvar <- ubar + (1 + 1 / m_eff) * Bvar
  se   <- sqrt(Tvar)
  rubin_df <- if (Bvar > 0) (m_eff - 1) * (1 + ubar / ((1 + 1 / m_eff) * Bvar))^2 else Inf
  if (!is.finite(rubin_df) || rubin_df <= 0) rubin_df <- Inf
  t_val <- qbar / se
  log("OK: pooled m_eff=", m_eff, "/", m,
      "; psi=", round(qbar, 4), "; se=", round(se, 4),
      "; df=", round(rubin_df, 1))
  list(
    psi = qbar, se = se, df = rubin_df, m = m_eff, m_attempted = m,
    n_per_fit = ns,
    p.value = 2 * pt(-abs(t_val), df = rubin_df),
    conf.low = qbar - qt(0.975, df = rubin_df) * se,
    conf.high = qbar + qt(0.975, df = rubin_df) * se,
    fits = fits
  )
}
svy_q_strict <- if (nrow(qdat_strict) > 0) {
  qdat_strict_svy <- as.data.frame(qdat_strict)
  for (v in c("phenoage_accel", "mixture_qscore", "ckd", "age", "pir", "bmi", "weight_combined", "psu", "strata")) {
    if (v %in% names(qdat_strict_svy)) qdat_strict_svy[[v]] <- suppressWarnings(as.numeric(as.character(qdat_strict_svy[[v]])))
  }
  qdat_strict_svy$ckd <- as.integer(qdat_strict_svy$ckd)
  qdat_strict_svy$sex <- droplevels(factor(qdat_strict_svy$sex))
  qdat_strict_svy$race_eth <- droplevels(factor(qdat_strict_svy$race_eth))
  qdat_strict_svy$cycle <- droplevels(factor(qdat_strict_svy$cycle))
  if ("smoker3"  %in% names(qdat_strict_svy)) qdat_strict_svy$smoker3  <- droplevels(factor(as.character(qdat_strict_svy$smoker3),  levels = c("Never", "Former", "Current")))
  if ("alcohol3" %in% names(qdat_strict_svy)) qdat_strict_svy$alcohol3 <- droplevels(factor(as.character(qdat_strict_svy$alcohol3), levels = c("Never", "Former", "Current")))
  tryCatch(
    svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined, nest = TRUE, data = qdat_strict_svy),
    error = function(e) svydesign(id = ~1, strata = NULL, weights = ~weight_combined, data = qdat_strict_svy)
  )
} else {
  NULL
}

# Model B: outcome is phenoage_accel (already age-residualized); do not adjust for age again (no double-dip).
covars_int <- c(
  if (valid_cat("sex", qdat)) "factor(sex)",
  if (valid_cat("race_eth", qdat)) "factor(race_eth)",
  if (valid_num("pir", qdat)) "pir",
  if (valid_cat("smoker3", qdat)) "factor(smoker3)",
  if (valid_num("bmi", qdat)) "bmi",
  if (valid_cat("alcohol3", qdat)) "factor(alcohol3)",
  if (valid_cat("cycle", qdat)) "factor(cycle)"
)
intended_covars_int <- covars_int
build_int_formula <- function(covars_vec) {
  rhs <- if (length(covars_vec) > 0) paste("+", paste(covars_vec, collapse = " + ")) else ""
  as.formula(paste("phenoage_accel ~ mixture_qscore * ckd", rhs))
}
f_int <- build_int_formula(covars_int)

repeat {
  mf_int <- tryCatch(model.frame(f_int, data = qdat, na.action = na.omit), error = function(e) NULL)
  if (is.null(mf_int) || nrow(mf_int) == 0) {
    stop("Model B preprocessing failed: no complete-case rows available for interaction formula.")
  }
  factor_terms <- names(mf_int)[vapply(mf_int, is.factor, logical(1))]
  bad_factor_terms <- factor_terms[vapply(mf_int[factor_terms], function(x) nlevels(droplevels(x)) < 2, logical(1))]
  removable_terms <- intersect(covars_int, bad_factor_terms)
  if (length(removable_terms) == 0) break
  message("Model B: dropping collapsed factor term(s) after complete-case filtering: ",
          paste(removable_terms, collapse = ", "))
  covars_int <- setdiff(covars_int, removable_terms)
  f_int <- build_int_formula(covars_int)
}
n_modelb <- nrow(mf_int)
realized_covars_int <- covars_int
covar_report <- tibble(
  intended_term = intended_covars_int,
  realized_in_model_b = ifelse(intended_covars_int %in% realized_covars_int, "Yes", "No")
)
kable(covar_report, caption = "Model B covariate transparency: intended vs realized adjustment terms.")

# ---- Model B fit: complete-case (always) AND MI-pooled (when use_mice = TRUE) ----
fit_int_cc_error <- NULL
fit_int_cc_raw <- tryCatch(
  svyglm(f_int, design = svy_q, family = gaussian()),
  error = function(e) { fit_int_cc_error <<- conditionMessage(e); NULL }
)
if (is.null(fit_int_cc_raw)) stop("Model B complete-case (svyglm interaction) failed: ", fit_int_cc_error)
fit_int_cc <- list(fit = fit_int_cc_raw, mi = FALSE)
message("Model B complete-case fit OK. N = ", n_modelb)
print(coef(fit_int_cc_raw))

fit_int_mi_error <- NULL
fit_int_mi <- if (mice_active) {
  tryCatch(
    fit_svyglm_pool(f_int, svy_q_mi),
    error = function(e) { fit_int_mi_error <<- conditionMessage(e); NULL }
  )
} else NULL
if (mice_active && is.null(fit_int_mi)) {
  message("Model B MI pooling failed: ", fit_int_mi_error, ". Falling back to complete-case.")
}

# Primary `fit_int` exposed downstream: MI-pooled if available, else complete-case.
fit_int <- if (!is.null(fit_int_mi)) fit_int_mi else fit_int_cc
n_modelb_primary <- if (!is.null(fit_int_mi)) n_mi_base else n_modelb
modelb_primary_label <- if (!is.null(fit_int_mi)) "MI-pooled (Rubin)" else "Complete-case"
message("Model B primary inference: ", modelb_primary_label, "; N = ", n_modelb_primary,
        " (complete-case N = ", n_modelb, ")")
# Sanity check: weighted lm and svyglm (weights-only design) should match (uses complete-case data).
fit_int_lm <- lm(f_int, data = qdat_svy, weights = weight_combined)
message("Weighted lm mixture_qscore coef: ", round(coef(fit_int_lm)["mixture_qscore"], 6),
        "; svyglm CC: ", round(coef(fit_int_cc_raw)["mixture_qscore"], 6))

aqg0 <- analytic_qg %>% filter(ckd == 0)
aqg1 <- analytic_qg %>% filter(ckd == 1)
if (nrow(aqg0) > 0) aqg0$weight_normalized <- aqg0$weight_combined / mean(aqg0$weight_combined, na.rm = TRUE)
if (nrow(aqg1) > 0) aqg1$weight_normalized <- aqg1$weight_combined / mean(aqg1$weight_combined, na.rm = TRUE)
fit_qg_ckd0 <- tryCatch(
  qgcomp::qgcomp.noboot(
    f_qg, expnms = qg_exposure_vars,
    data = aqg0,
    family = gaussian(), q = q_qg,
    weights = aqg0$weight_normalized
  ),
  error = function(e) NULL
)
fit_qg_ckd1 <- tryCatch(
  qgcomp::qgcomp.noboot(
    f_qg, expnms = qg_exposure_vars,
    data = aqg1,
    family = gaussian(), q = q_qg,
    weights = aqg1$weight_normalized
  ),
  error = function(e) NULL
)
```

# Sample Flow (PRISMA-style)

```{r sample-flow}
n0 <- nrow(nhanes_raw)
n1 <- sum(dat$age >= config$min_age, na.rm = TRUE)
# Split the old "PhenoAge AND CKD non-missing" step into two consecutive exclusions:
# (a) missing PhenoAge components -> phenoage is NA when any required biomarker is missing;
# (b) missing CKD modelling components -> ckd is NA when both eGFR and ACR cannot be computed.
n2a <- sum(dat$age >= config$min_age & !is.na(dat$phenoage), na.rm = TRUE)
n2b <- sum(dat$age >= config$min_age & !is.na(dat$phenoage) & !is.na(dat$ckd), na.rm = TRUE)
n3 <- nrow(analytic)
n4 <- if (exists("qdat")) nrow(qdat) else NA_integer_
n5 <- if (exists("analytic_qg")) nrow(analytic_qg) else NA_integer_
n6 <- if (exists("qdat_strict")) nrow(qdat_strict) else NA_integer_
n7 <- if (exists("n_modelb")) n_modelb else NA_integer_
n8 <- if (exists("n_modelb_primary")) n_modelb_primary else NA_integer_
mi_label_step <- if (exists("mice_active") && isTRUE(mice_active)) {
  paste0("Model B PRIMARY estimation sample (MICE m=", config$mice_m, "; covariates imputed)")
} else {
  "Model B PRIMARY estimation sample (complete-case; MICE off)"
}

flow <- tibble(
  step = c(
    "Downloaded NHANES participants (selected cycles)",
    paste0("Age >= ", config$min_age),
    "Non-missing PhenoAge components (all 9 biomarkers required)",
    "Non-missing CKD modelling components (eGFR or ACR available)",
    "Non-missing selected weight, PSU, and strata (analytic sample)",
    "Primary interaction model sample (fixed 4-PFAS complete-case score + design vars)",
    "Complete-case qgcomp sample",
    "Strict complete-case (qgcomp-compatible PFAS subset)",
    "Model B complete-case estimation sample (covariate complete-case)",
    mi_label_step
  ),
  included = c(n0, n1, n2a, n2b, n3, n4, n5, n6, n7, n8)
) %>%
  mutate(
    excluded_at_step = pmax(lag(included, default = first(included)) - included, 0),
    excluded_at_step = if_else(row_number() == 1, NA_integer_, as.integer(excluded_at_step)),
    excluded_label = ifelse(is.na(excluded_at_step), "", paste0("Excluded: ", excluded_at_step))
  )

kable(flow, caption = "Sample flow counts (PRISMA-style): inclusion and exclusion at each analytic step")

flow_plot <- flow %>%
  mutate(
    y = rev(seq_len(n())),
    box_label = paste0(step, "\nN = ", included)
  )

ggplot(flow_plot, aes(x = 1, y = y)) +
  geom_segment(
    data = flow_plot %>% filter(y > min(y)),
    aes(x = 1, xend = 1, y = y - 0.42, yend = y - 0.95),
    arrow = arrow(length = unit(0.12, "inches")),
    linewidth = 0.5
  ) +
  geom_label(aes(label = box_label), size = 3.2, label.size = 0.2, fill = "white") +
  geom_text(aes(x = 1.62, label = excluded_label), hjust = 0, size = 3.1, color = "gray30") +
  xlim(0.5, 2.2) +
  ylim(0.5, max(flow_plot$y) + 0.8) +
  labs(
    title = "Participant flow from downloaded sample to analytic samples",
    subtitle = "PRISMA-style flow for NHANES analytic cohort development",
    x = NULL, y = NULL
  ) +
  theme_void(base_size = 11)
```

# PFAS Exposure Distributions

```{r pfas-descriptives}
# Raw serum PFAS (ng/mL) for the four primary chemicals used in the mixture.
pfas_raw_map <- c(
  "PFOS"  = "LBXPFOS",
  "PFOA"  = "LBXPFOA",
  "PFNA"  = "LBXPFNA",
  "PFHxS" = "LBXPFHS"
)
pfas_present <- pfas_raw_map[pfas_raw_map %in% names(analytic)]

# Wide -> long for plotting (one panel per chemical) using the analytic sample.
# Coerce all PFAS columns to numeric first (some NHANES vars arrive as character).
pfas_wide <- analytic %>%
  dplyr::select(dplyr::any_of(unname(pfas_present))) %>%
  dplyr::mutate(dplyr::across(dplyr::everything(), function(x) {
    suppressWarnings(as.numeric(trimws(as.character(x))))
  }))

pfas_long <- pfas_wide %>%
  tidyr::pivot_longer(dplyr::everything(), names_to = "var", values_to = "value") %>%
  dplyr::filter(!is.na(value), value > 0) %>%
  dplyr::mutate(chemical = factor(names(pfas_present)[match(var, pfas_present)],
                                  levels = names(pfas_present)))

# Geometric mean / SD helpers (applied on positive values only)
gm_mean <- function(x) {
  x <- x[is.finite(x) & x > 0]
  if (!length(x)) NA_real_ else exp(mean(log(x)))
}
gm_sd <- function(x) {
  x <- x[is.finite(x) & x > 0]
  if (length(x) < 2) NA_real_ else exp(stats::sd(log(x)))
}

pfas_desc <- pfas_long %>%
  dplyr::group_by(Chemical = chemical) %>%
  dplyr::summarise(
    N        = sum(!is.na(value)),
    Mean     = mean(value, na.rm = TRUE),
    SD       = stats::sd(value, na.rm = TRUE),
    Median   = stats::median(value, na.rm = TRUE),
    Q1       = stats::quantile(value, 0.25, na.rm = TRUE),
    Q3       = stats::quantile(value, 0.75, na.rm = TRUE),
    IQR      = Q3 - Q1,
    Min      = min(value, na.rm = TRUE),
    Max      = max(value, na.rm = TRUE),
    GeoMean  = gm_mean(value),
    GeoSD    = gm_sd(value),
    .groups  = "drop"
  ) %>%
  dplyr::mutate(dplyr::across(c(Mean, SD, Median, Q1, Q3, IQR, Min, Max, GeoMean, GeoSD),
                              ~ round(.x, 3)))

knitr::kable(
  pfas_desc,
  caption = "PFAS descriptive statistics in the analytic sample (serum, ng/mL). Geometric mean / SD computed on positive values only."
)

# Histogram panels (2x2): export both linear axis (true raw scale) and log10-axis (skewness readability).
hist_base <- function() {
  ggplot(pfas_long, aes(x = value)) +
    geom_histogram(bins = 40, fill = "grey80", color = "black", linewidth = 0.2) +
    facet_wrap(~ chemical, ncol = 2, scales = "free") +
    theme_bw(base_size = 10) +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, linewidth = 0.4),
          strip.background = element_blank(),
          strip.text = element_text(face = "bold"))
}
pfas_hist_linear <- hist_base() +
  labs(x = "Serum concentration (ng/mL)", y = "Count",
       title = "PFAS distributions (linear x-axis)")
pfas_hist_log10 <- hist_base() +
  scale_x_log10() +
  labs(x = "Serum concentration (ng/mL, log10 axis)", y = "Count",
       title = "PFAS distributions (log10-scaled axis)")

dir.create("figures", showWarnings = FALSE)
ggsave("figures/fig0_pfas_distributions_linear.jpg", plot = pfas_hist_linear,
       width = 8.5, height = 5.5, dpi = 300, device = "jpeg",
       quality = 95, bg = "white")
ggsave("figures/fig0_pfas_distributions_log10.jpg", plot = pfas_hist_log10,
       width = 8.5, height = 5.5, dpi = 300, device = "jpeg",
       quality = 95, bg = "white")
# Backwards-compatible alias (linear = raw concentrations)
file.copy("figures/fig0_pfas_distributions_linear.jpg",
          "figures/fig0_pfas_distributions.jpg", overwrite = TRUE)

pfas_hist_linear / pfas_hist_log10 +
  plot_layout(ncol = 1) &
  theme(plot.title = element_text(face = "bold", size = 11))
```

# Tables

```{r tables}
fmt <- function(x, digits = 3) {
  ifelse(
    is.na(x),
    "NA",
    ifelse(
      x != 0 & abs(x) < 10^-digits,
      format(x, scientific = TRUE, digits = 2),
      format(round(x, digits), nsmall = digits)
    )
  )
}
fmt_p <- function(x) ifelse(is.na(x), "NA", ifelse(x < 0.001, "<0.001", format(round(x, 3), nsmall = 3)))

# Table 1: characteristics by CKD -- continuous: mean (SD); categorical: n (weighted %)
t1_mean_sd <- function(formula, design) {
  est <- svymean(formula, design, na.rm = TRUE)
  v_obj <- svyvar(formula, design, na.rm = TRUE)
  sd_val <- sqrt(as.numeric(coef(v_obj)[1]))
  paste0(format(round(coef(est)[1], 2), nsmall = 2), " (", format(round(sd_val, 2), nsmall = 2), ")")
}
t1_mean_obj <- function(formula, design) {
  m <- svymean(formula, design, na.rm = TRUE)
  v <- svyvar(formula, design, na.rm = TRUE)
  list(est = unname(coef(m)[1]), se = unname(SE(m)[1]), var = unname(coef(v)[1]))
}
t1_pct_se <- function(formula, design) {
  est <- svymean(formula, design, na.rm = TRUE)
  cf <- coef(est)
  se <- SE(est)
  idx <- 1
  if (length(cf) > 1) {
    nms <- names(cf)
    idx_true <- which(str_detect(nms, "TRUE$"))
    if (length(idx_true) == 0) idx_true <- which(str_detect(nms, "1$"))
    if (length(idx_true) > 0) idx <- idx_true[1]
  }
  paste0(
    format(round(unname(cf[idx]) * 100, 1), nsmall = 1),
    " (",
    format(round(unname(se[idx]) * 100, 1), nsmall = 1),
    ")"
  )
}
t1_n_wpct <- function(cond_vec, formula, design) {
  n_unwt <- sum(cond_vec, na.rm = TRUE)
  est <- svymean(formula, design, na.rm = TRUE)
  cf <- coef(est)
  idx <- 1
  if (length(cf) > 1) {
    nms <- names(cf)
    idx_true <- which(str_detect(nms, "TRUE$"))
    if (length(idx_true) == 0) idx_true <- which(str_detect(nms, "1$"))
    if (length(idx_true) > 0) idx <- idx_true[1]
  }
  paste0(
    format(n_unwt, big.mark = ","),
    " (",
    format(round(unname(cf[idx]) * 100, 1), nsmall = 1),
    ")"
  )
}
t1_pct_obj <- function(formula, design) {
  est <- svymean(formula, design, na.rm = TRUE)
  cf <- coef(est)
  se <- SE(est)
  idx <- 1
  if (length(cf) > 1) {
    nms <- names(cf)
    idx_true <- which(str_detect(nms, "TRUE$"))
    if (length(idx_true) == 0) idx_true <- which(str_detect(nms, "1$"))
    if (length(idx_true) > 0) idx <- idx_true[1]
  }
  list(est = unname(cf[idx]), se = unname(se[idx]))
}
t1_factor_level_pct_se <- function(var, level, design) {
  est <- svymean(as.formula(paste0("~factor(", var, ")")), design, na.rm = TRUE)
  key <- paste0("factor(", var, ")", level)
  if (!(key %in% names(coef(est)))) return("NA")
  paste0(
    format(round(unname(coef(est)[key]) * 100, 1), nsmall = 1),
    " (",
    format(round(unname(SE(est)[key]) * 100, 1), nsmall = 1),
    ")"
  )
}
t1_factor_level_n_wpct <- function(var, level, design, data) {
  n_unwt <- sum(data[[var]] == level, na.rm = TRUE)
  est <- svymean(as.formula(paste0("~factor(", var, ")")), design, na.rm = TRUE)
  key <- paste0("factor(", var, ")", level)
  if (!(key %in% names(coef(est)))) {
    return(paste0(format(n_unwt, big.mark = ","), " (NA)"))
  }
  paste0(
    format(n_unwt, big.mark = ","),
    " (",
    format(round(unname(coef(est)[key]) * 100, 1), nsmall = 1),
    ")"
  )
}
t1_factor_level_obj <- function(var, level, design) {
  est <- svymean(as.formula(paste0("~factor(", var, ")")), design, na.rm = TRUE)
  key <- paste0("factor(", var, ")", level)
  if (!(key %in% names(coef(est)))) return(list(est = NA_real_, se = NA_real_))
  list(est = unname(coef(est)[key]), se = unname(SE(est)[key]))
}
# Missing category: weighted % of rows with NA on `var` (for MICE / missingness diagnostics).
t1_miss_pct_obj <- function(var, design) {
  t1_pct_obj(as.formula(paste0("~I(is.na(", var, "))")), design)
}
t1_miss_n_wpct <- function(var, design, data_sub) {
  n_unwt <- sum(is.na(data_sub[[var]]))
  o <- t1_miss_pct_obj(var, design)
  paste0(
    format(n_unwt, big.mark = ","),
    " (",
    format(round(o$est * 100, 1), nsmall = 1),
    ")"
  )
}
smd_cont <- function(o0, o1) {
  if (any(!is.finite(c(o0$est, o1$est, o0$var, o1$var)))) return(NA_real_)
  denom <- sqrt((o0$var + o1$var) / 2)
  if (!is.finite(denom) || denom <= 0) return(NA_real_)
  abs(o1$est - o0$est) / denom
}
smd_prop <- function(o0, o1) {
  if (any(!is.finite(c(o0$est, o1$est)))) return(NA_real_)
  denom <- sqrt((o0$est * (1 - o0$est) + o1$est * (1 - o1$est)) / 2)
  if (!is.finite(denom) || denom <= 0) return(NA_real_)
  abs(o1$est - o0$est) / denom
}
fmt_smd <- function(x) ifelse(is.na(x), "", format(round(x, 3), nsmall = 3))

# Build Part A (Model B PRIMARY estimation sample = MICE base; PFAS+outcome+design complete) and
# Part B (analytic-sample participants excluded from primary inference: missing PFAS / outcome / design).
# Within Part A we display Missing rows so the reader sees exactly which covariate cells are imputed by MICE.
if ("SEQN" %in% names(analytic) && exists("qdat_svy") && "SEQN" %in% names(qdat_svy)) {
  inc_idx <- analytic$SEQN %in% qdat_svy$SEQN
} else {
  inc_idx <- !is.na(analytic$ckd) & !is.na(analytic$phenoage_accel)
}
analytic_inc <- analytic[inc_idx, , drop = FALSE]
analytic_exc <- analytic[!inc_idx, , drop = FALSE]

svy_inc <- svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined,
                     nest = TRUE, data = analytic_inc)
svy_exc <- svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined,
                     nest = TRUE, data = analytic_exc)

# Block builder: returns a tibble with columns characteristic, No CKD, CKD, SMD.
# Continuous: mean (SD).  Categorical: n (weighted %).  No SE displayed anywhere.
build_t1_block <- function(design, data) {
  d0 <- subset(design, ckd == 0)
  d1 <- subset(design, ckd == 1)
  data0 <- data[!is.na(data$ckd) & data$ckd == 0, , drop = FALSE]
  data1 <- data[!is.na(data$ckd) & data$ckd == 1, , drop = FALSE]

  t1_show_miss <- function(var) var %in% names(data) && any(is.na(data[[var]]))

  age0 <- t1_mean_obj(~age, d0); age1 <- t1_mean_obj(~age, d1)
  age_miss0 <- t1_miss_pct_obj("age", d0); age_miss1 <- t1_miss_pct_obj("age", d1)
  pir0 <- t1_mean_obj(~pir, d0); pir1 <- t1_mean_obj(~pir, d1)
  pir_miss0 <- t1_miss_pct_obj("pir", d0); pir_miss1 <- t1_miss_pct_obj("pir", d1)
  bmi0 <- t1_mean_obj(~bmi, d0); bmi1 <- t1_mean_obj(~bmi, d1)
  bmi_miss0 <- t1_miss_pct_obj("bmi", d0); bmi_miss1 <- t1_miss_pct_obj("bmi", d1)

  fem0 <- t1_pct_obj(~I(sex == "Female"), d0); fem1 <- t1_pct_obj(~I(sex == "Female"), d1)
  sex_miss0 <- t1_miss_pct_obj("sex", d0); sex_miss1 <- t1_miss_pct_obj("sex", d1)
  his0 <- t1_factor_level_obj("race_eth", "Hispanic", d0); his1 <- t1_factor_level_obj("race_eth", "Hispanic", d1)
  whi0 <- t1_factor_level_obj("race_eth", "Non-Hispanic White", d0); whi1 <- t1_factor_level_obj("race_eth", "Non-Hispanic White", d1)
  blk0 <- t1_factor_level_obj("race_eth", "Non-Hispanic Black", d0); blk1 <- t1_factor_level_obj("race_eth", "Non-Hispanic Black", d1)
  asn0 <- t1_factor_level_obj("race_eth", "Non-Hispanic Asian", d0); asn1 <- t1_factor_level_obj("race_eth", "Non-Hispanic Asian", d1)
  oth0 <- t1_factor_level_obj("race_eth", "Other/Multi", d0); oth1 <- t1_factor_level_obj("race_eth", "Other/Multi", d1)
  race_miss0 <- t1_miss_pct_obj("race_eth", d0); race_miss1 <- t1_miss_pct_obj("race_eth", d1)
  edu_lths0 <- t1_factor_level_obj("education", "<HS", d0); edu_lths1 <- t1_factor_level_obj("education", "<HS", d1)
  edu_hs0   <- t1_factor_level_obj("education", "HS/GED", d0); edu_hs1   <- t1_factor_level_obj("education", "HS/GED", d1)
  edu_gt0   <- t1_factor_level_obj("education", ">HS", d0); edu_gt1   <- t1_factor_level_obj("education", ">HS", d1)
  edu_miss0 <- t1_miss_pct_obj("education", d0); edu_miss1 <- t1_miss_pct_obj("education", d1)
  smk_never0 <- t1_factor_level_obj("smoker3", "Never",   d0); smk_never1 <- t1_factor_level_obj("smoker3", "Never",   d1)
  smk_form0  <- t1_factor_level_obj("smoker3", "Former",  d0); smk_form1  <- t1_factor_level_obj("smoker3", "Former",  d1)
  smk_curr0  <- t1_factor_level_obj("smoker3", "Current", d0); smk_curr1  <- t1_factor_level_obj("smoker3", "Current", d1)
  smk_miss0 <- t1_miss_pct_obj("smoker3", d0); smk_miss1 <- t1_miss_pct_obj("smoker3", d1)
  alc_never0 <- t1_factor_level_obj("alcohol3", "Never",   d0); alc_never1 <- t1_factor_level_obj("alcohol3", "Never",   d1)
  alc_form0  <- t1_factor_level_obj("alcohol3", "Former",  d0); alc_form1  <- t1_factor_level_obj("alcohol3", "Former",  d1)
  alc_curr0  <- t1_factor_level_obj("alcohol3", "Current", d0); alc_curr1  <- t1_factor_level_obj("alcohol3", "Current", d1)
  alc_miss0 <- t1_miss_pct_obj("alcohol3", d0); alc_miss1 <- t1_miss_pct_obj("alcohol3", d1)
  pa0  <- t1_pct_obj(~I(phys_active == 1), d0); pa1  <- t1_pct_obj(~I(phys_active == 1), d1)
  pa_miss0 <- t1_miss_pct_obj("phys_active", d0); pa_miss1 <- t1_miss_pct_obj("phys_active", d1)
  dm0  <- t1_pct_obj(~I(diabetes == 1), d0); dm1  <- t1_pct_obj(~I(diabetes == 1), d1)
  dm_miss0 <- t1_miss_pct_obj("diabetes", d0); dm_miss1 <- t1_miss_pct_obj("diabetes", d1)
  htn0 <- t1_pct_obj(~I(hypertension == 1), d0); htn1 <- t1_pct_obj(~I(hypertension == 1), d1)
  htn_miss0 <- t1_miss_pct_obj("hypertension", d0); htn_miss1 <- t1_miss_pct_obj("hypertension", d1)
  cvd0 <- t1_pct_obj(~I(cvd == 1), d0); cvd1 <- t1_pct_obj(~I(cvd == 1), d1)
  cvd_miss0 <- t1_miss_pct_obj("cvd", d0); cvd_miss1 <- t1_miss_pct_obj("cvd", d1)

  show_age_miss <- t1_show_miss("age")
  show_pir_miss <- t1_show_miss("pir")
  show_bmi_miss <- t1_show_miss("bmi")
  show_sex_miss <- t1_show_miss("sex")
  show_race_miss <- t1_show_miss("race_eth")
  show_edu_miss <- t1_show_miss("education")
  show_smk_miss <- t1_show_miss("smoker3")
  show_alc_miss <- t1_show_miss("alcohol3")
  show_pa_miss <- t1_show_miss("phys_active")
  show_dm_miss <- t1_show_miss("diabetes")
  show_htn_miss <- t1_show_miss("hypertension")
  show_cvd_miss <- t1_show_miss("cvd")

  tibble(
    characteristic = c(
      "N (unweighted)", "Age, mean (SD)",
      if (show_age_miss) "Age missing, n (weighted %)" else character(0),
      "Female, n (weighted %)",
      if (show_sex_miss) "Sex missing, n (weighted %)" else character(0),
      "Hispanic, n (weighted %)", "Non-Hispanic White, n (weighted %)", "Non-Hispanic Black, n (weighted %)",
      "Non-Hispanic Asian, n (weighted %)", "Other/Multi, n (weighted %)",
      if (show_race_miss) "Race/ethnicity missing, n (weighted %)" else character(0),
      "Education <HS, n (weighted %)", "Education HS/GED, n (weighted %)", "Education >HS, n (weighted %)",
      if (show_edu_miss) "Education missing, n (weighted %)" else character(0),
      "PIR, mean (SD)",
      if (show_pir_miss) "PIR missing, n (weighted %)" else character(0),
      "BMI, mean (SD)",
      if (show_bmi_miss) "BMI missing, n (weighted %)" else character(0),
      "Never smoker, n (weighted %)", "Former smoker, n (weighted %)", "Current smoker, n (weighted %)",
      if (show_smk_miss) "Smoking status missing, n (weighted %)" else character(0),
      "Never drinker, n (weighted %)", "Former drinker, n (weighted %)", "Current drinker, n (weighted %)",
      if (show_alc_miss) "Alcohol status missing, n (weighted %)" else character(0),
      "Physically active, n (weighted %)",
      if (show_pa_miss) "Physical activity missing, n (weighted %)" else character(0),
      "Diabetes, n (weighted %)",
      if (show_dm_miss) "Diabetes missing, n (weighted %)" else character(0),
      "Hypertension, n (weighted %)",
      if (show_htn_miss) "Hypertension missing, n (weighted %)" else character(0),
      "CVD history, n (weighted %)",
      if (show_cvd_miss) "CVD history missing, n (weighted %)" else character(0)
    ),
    `No CKD` = c(
      format(nrow(data0), big.mark = ","),
      t1_mean_sd(~age, d0),
      if (show_age_miss) t1_miss_n_wpct("age", d0, data0) else character(0),
      t1_n_wpct(data0$sex == "Female", ~I(sex == "Female"), d0),
      if (show_sex_miss) t1_miss_n_wpct("sex", d0, data0) else character(0),
      t1_factor_level_n_wpct("race_eth", "Hispanic", d0, data0),
      t1_factor_level_n_wpct("race_eth", "Non-Hispanic White", d0, data0),
      t1_factor_level_n_wpct("race_eth", "Non-Hispanic Black", d0, data0),
      t1_factor_level_n_wpct("race_eth", "Non-Hispanic Asian", d0, data0),
      t1_factor_level_n_wpct("race_eth", "Other/Multi", d0, data0),
      if (show_race_miss) t1_miss_n_wpct("race_eth", d0, data0) else character(0),
      t1_factor_level_n_wpct("education", "<HS", d0, data0),
      t1_factor_level_n_wpct("education", "HS/GED", d0, data0),
      t1_factor_level_n_wpct("education", ">HS", d0, data0),
      if (show_edu_miss) t1_miss_n_wpct("education", d0, data0) else character(0),
      t1_mean_sd(~pir, d0),
      if (show_pir_miss) t1_miss_n_wpct("pir", d0, data0) else character(0),
      t1_mean_sd(~bmi, d0),
      if (show_bmi_miss) t1_miss_n_wpct("bmi", d0, data0) else character(0),
      t1_factor_level_n_wpct("smoker3", "Never",   d0, data0),
      t1_factor_level_n_wpct("smoker3", "Former",  d0, data0),
      t1_factor_level_n_wpct("smoker3", "Current", d0, data0),
      if (show_smk_miss) t1_miss_n_wpct("smoker3", d0, data0) else character(0),
      t1_factor_level_n_wpct("alcohol3", "Never",   d0, data0),
      t1_factor_level_n_wpct("alcohol3", "Former",  d0, data0),
      t1_factor_level_n_wpct("alcohol3", "Current", d0, data0),
      if (show_alc_miss) t1_miss_n_wpct("alcohol3", d0, data0) else character(0),
      t1_n_wpct(data0$phys_active == 1, ~I(phys_active == 1), d0),
      if (show_pa_miss) t1_miss_n_wpct("phys_active", d0, data0) else character(0),
      t1_n_wpct(data0$diabetes == 1, ~I(diabetes == 1), d0),
      if (show_dm_miss) t1_miss_n_wpct("diabetes", d0, data0) else character(0),
      t1_n_wpct(data0$hypertension == 1, ~I(hypertension == 1), d0),
      if (show_htn_miss) t1_miss_n_wpct("hypertension", d0, data0) else character(0),
      t1_n_wpct(data0$cvd == 1, ~I(cvd == 1), d0),
      if (show_cvd_miss) t1_miss_n_wpct("cvd", d0, data0) else character(0)
    ),
    CKD = c(
      format(nrow(data1), big.mark = ","),
      t1_mean_sd(~age, d1),
      if (show_age_miss) t1_miss_n_wpct("age", d1, data1) else character(0),
      t1_n_wpct(data1$sex == "Female", ~I(sex == "Female"), d1),
      if (show_sex_miss) t1_miss_n_wpct("sex", d1, data1) else character(0),
      t1_factor_level_n_wpct("race_eth", "Hispanic", d1, data1),
      t1_factor_level_n_wpct("race_eth", "Non-Hispanic White", d1, data1),
      t1_factor_level_n_wpct("race_eth", "Non-Hispanic Black", d1, data1),
      t1_factor_level_n_wpct("race_eth", "Non-Hispanic Asian", d1, data1),
      t1_factor_level_n_wpct("race_eth", "Other/Multi", d1, data1),
      if (show_race_miss) t1_miss_n_wpct("race_eth", d1, data1) else character(0),
      t1_factor_level_n_wpct("education", "<HS", d1, data1),
      t1_factor_level_n_wpct("education", "HS/GED", d1, data1),
      t1_factor_level_n_wpct("education", ">HS", d1, data1),
      if (show_edu_miss) t1_miss_n_wpct("education", d1, data1) else character(0),
      t1_mean_sd(~pir, d1),
      if (show_pir_miss) t1_miss_n_wpct("pir", d1, data1) else character(0),
      t1_mean_sd(~bmi, d1),
      if (show_bmi_miss) t1_miss_n_wpct("bmi", d1, data1) else character(0),
      t1_factor_level_n_wpct("smoker3", "Never",   d1, data1),
      t1_factor_level_n_wpct("smoker3", "Former",  d1, data1),
      t1_factor_level_n_wpct("smoker3", "Current", d1, data1),
      if (show_smk_miss) t1_miss_n_wpct("smoker3", d1, data1) else character(0),
      t1_factor_level_n_wpct("alcohol3", "Never",   d1, data1),
      t1_factor_level_n_wpct("alcohol3", "Former",  d1, data1),
      t1_factor_level_n_wpct("alcohol3", "Current", d1, data1),
      if (show_alc_miss) t1_miss_n_wpct("alcohol3", d1, data1) else character(0),
      t1_n_wpct(data1$phys_active == 1, ~I(phys_active == 1), d1),
      if (show_pa_miss) t1_miss_n_wpct("phys_active", d1, data1) else character(0),
      t1_n_wpct(data1$diabetes == 1, ~I(diabetes == 1), d1),
      if (show_dm_miss) t1_miss_n_wpct("diabetes", d1, data1) else character(0),
      t1_n_wpct(data1$hypertension == 1, ~I(hypertension == 1), d1),
      if (show_htn_miss) t1_miss_n_wpct("hypertension", d1, data1) else character(0),
      t1_n_wpct(data1$cvd == 1, ~I(cvd == 1), d1),
      if (show_cvd_miss) t1_miss_n_wpct("cvd", d1, data1) else character(0)
    ),
    SMD = c(
      "",
      fmt_smd(smd_cont(age0, age1)),
      if (show_age_miss) fmt_smd(smd_prop(age_miss0, age_miss1)) else character(0),
      fmt_smd(smd_prop(fem0, fem1)),
      if (show_sex_miss) fmt_smd(smd_prop(sex_miss0, sex_miss1)) else character(0),
      fmt_smd(smd_prop(his0, his1)),
      fmt_smd(smd_prop(whi0, whi1)),
      fmt_smd(smd_prop(blk0, blk1)),
      fmt_smd(smd_prop(asn0, asn1)),
      fmt_smd(smd_prop(oth0, oth1)),
      if (show_race_miss) fmt_smd(smd_prop(race_miss0, race_miss1)) else character(0),
      fmt_smd(smd_prop(edu_lths0, edu_lths1)),
      fmt_smd(smd_prop(edu_hs0, edu_hs1)),
      fmt_smd(smd_prop(edu_gt0, edu_gt1)),
      if (show_edu_miss) fmt_smd(smd_prop(edu_miss0, edu_miss1)) else character(0),
      fmt_smd(smd_cont(pir0, pir1)),
      if (show_pir_miss) fmt_smd(smd_prop(pir_miss0, pir_miss1)) else character(0),
      fmt_smd(smd_cont(bmi0, bmi1)),
      if (show_bmi_miss) fmt_smd(smd_prop(bmi_miss0, bmi_miss1)) else character(0),
      fmt_smd(smd_prop(smk_never0, smk_never1)),
      fmt_smd(smd_prop(smk_form0,  smk_form1)),
      fmt_smd(smd_prop(smk_curr0,  smk_curr1)),
      if (show_smk_miss) fmt_smd(smd_prop(smk_miss0, smk_miss1)) else character(0),
      fmt_smd(smd_prop(alc_never0, alc_never1)),
      fmt_smd(smd_prop(alc_form0,  alc_form1)),
      fmt_smd(smd_prop(alc_curr0,  alc_curr1)),
      if (show_alc_miss) fmt_smd(smd_prop(alc_miss0, alc_miss1)) else character(0),
      fmt_smd(smd_prop(pa0, pa1)),
      if (show_pa_miss) fmt_smd(smd_prop(pa_miss0, pa_miss1)) else character(0),
      fmt_smd(smd_prop(dm0, dm1)),
      if (show_dm_miss) fmt_smd(smd_prop(dm_miss0, dm_miss1)) else character(0),
      fmt_smd(smd_prop(htn0, htn1)),
      if (show_htn_miss) fmt_smd(smd_prop(htn_miss0, htn_miss1)) else character(0),
      fmt_smd(smd_prop(cvd0, cvd1)),
      if (show_cvd_miss) fmt_smd(smd_prop(cvd_miss0, cvd_miss1)) else character(0)
    )
  )
}

block_a <- build_t1_block(svy_inc, analytic_inc)
block_b <- build_t1_block(svy_exc, analytic_exc)

n_inc <- nrow(analytic_inc)
n_exc <- nrow(analytic_exc)
n_inc_0 <- sum(analytic_inc$ckd == 0, na.rm = TRUE)
n_inc_1 <- sum(analytic_inc$ckd == 1, na.rm = TRUE)
n_exc_0 <- sum(analytic_exc$ckd == 0, na.rm = TRUE)
n_exc_1 <- sum(analytic_exc$ckd == 1, na.rm = TRUE)

header_a <- tibble(
  characteristic = "PART A -- INCLUDED IN STUDY",
  `No CKD` = "", CKD = "", SMD = ""
)
header_b <- tibble(
  characteristic = "PART B -- EXCLUDED FROM STUDY",
  `No CKD` = "", CKD = "", SMD = ""
)

table1 <- bind_rows(header_a, block_a, header_b, block_b)

kable(
  table1,
  caption = "Table 1. Sample characteristics by CKD status. Part A = MICE base sample (Model B PRIMARY estimation sample; PFAS + outcome + design complete). Missing rows show pre-imputation missingness on each variable shown (continuous and categorical alike); those cells are then imputed under MICE for primary inference. Continuous: mean (SD); categorical: n (weighted %); Missing rows: n (weighted %) of NA. SMD is the absolute standardized mean difference between CKD vs No CKD within each part."
)
cat("\n*Table 1 note:* Weighted percentages use the NHANES complex survey design (strata, PSU, combined weights). ",
    "Part A = MICE base sample = participants entering Model B primary inference (non-missing PhenoAge, CKD, weight/PSU/strata, and complete primary 4-PFAS set). ",
    "Within Part A, the Missing rows quantify pre-imputation covariate missingness; those cells are imputed via MICE (m = ", config$mice_m,
    ") and pooled with Rubin's rules. ",
    "Part B = analytic-sample participants excluded from primary inference because they are missing one or more PFAS / outcome / design components (these are NOT imputed). ",
    "SMD in Part B is computed between excluded No CKD vs excluded CKD; values >0.10 are typically considered meaningful imbalance.\n", sep = "")

# Table 2: overall mixture effect (MI-pooled qgcomp via Rubin when MICE is active; CC fallback otherwise).
qg_mi_overall <- if (mice_active) {
  qgcomp_mi_pool(
    formula = f_qg, expnms = qg_exposure_vars,
    imp_list_obj = imp_list, q = q_qg,
    weight_col = "weight_combined"
  )
} else NULL

tab2_mi <- if (!is.null(qg_mi_overall)) {
  tibble(
    model = paste0("Model A: Overall mixture effect [MI-pooled, m = ", qg_mi_overall$m, "]"),
    beta = fmt(qg_mi_overall$psi),
    conf.low = fmt(qg_mi_overall$conf.low),
    conf.high = fmt(qg_mi_overall$conf.high),
    p.value = fmt_p(qg_mi_overall$p.value),
    N = round(mean(qg_mi_overall$n_per_fit)),
    source = "MI-pooled"
  )
} else NULL

tab2_cc <- safe_qgcomp_extract(
  fit_qg_overall,
  if (!is.null(qg_mi_overall)) "Model A: Overall mixture effect [Complete-case sensitivity]"
  else "Model A: Overall mixture effect [Complete-case]",
  nrow(analytic_qg)
) %>%
  mutate(
    beta = fmt(beta),
    conf.low = fmt(conf.low),
    conf.high = fmt(conf.high),
    p.value = fmt_p(p.value),
    source = "Complete-case"
  )

tab2_full <- if (!is.null(tab2_mi)) bind_rows(tab2_mi, tab2_cc) else tab2_cc

kable(
  tab2_full,
  caption = paste0(
    "Table 2. Overall association between PFAS mixture score and PhenoAge acceleration (years). ",
    "Mixture score ranges from 1 to 4; a 1-unit increase represents a 1-quantile increase averaged ",
    "across the fixed primary PFAS set. Estimates from qgcomp with normalized pooled NHANES sampling weights",
    if (!is.null(qg_mi_overall)) "; primary row is MI-pooled across imputations via Rubin's rules; CC row shown for transparency." else " (complete-case)."
  )
)
cat("\n*Table 2 note:* qgcomp model adjusted for the same covariates as Model B: sex, race/ethnicity, PIR, BMI, smoking (Never/Former/Current), alcohol (Never/Former/Current), cycle (no age; outcome is age-residualized). ",
    "qgcomp intended covariates: ", paste(qg_covars_intended, collapse = ", "), ". ",
    "qgcomp realized covariates (after complete-case support checks): ", paste(qg_covars_realized, collapse = ", "), ". ",
    if (!is.null(qg_mi_overall)) {
      paste0("Primary row uses ", qg_mi_overall$m, " MICE imputations of model covariates; psi (overall mixture coef) and its variance are pooled via Rubin's rules. ",
             "qgcomp variance does NOT incorporate PSU/strata; statistical inference for the interaction question relies on the survey-weighted Table 3 (svyglm). ")
    } else if (!is.null(fit_qg_overall)) {
      "Inference here uses qgcomp's analytic variance on the complete-case sample; see Table 3 (svyglm) for primary survey-weighted inference. "
    } else {
      paste0("qgcomp failed: ",
             ifelse(is.null(qgcomp_overall_error), "unknown error", qgcomp_overall_error), ". ")
    },
    "PhenoAge acceleration defined as survey-weighted residual of PhenoAge on chronological age.\n", sep = "")

# Table 3: interaction results + design-based marginal effects by CKD group
if (!is.null(fit_int)) {
  int_tidy <- tidy_pool(fit_int, conf.level = 0.95)
  cf_names <- names(coef_pool(fit_int))

  int_terms_wanted <- c("mixture_qscore", cf_names[str_detect(cf_names, "^ckd$|mixture_qscore:ckd")])
  tab3_raw <- int_tidy %>%
    filter(term %in% int_terms_wanted) %>%
    transmute(
      term,
      beta = fmt(estimate),
      conf.low = fmt(conf.low),
      conf.high = fmt(conf.high),
      p.value = fmt_p(p.value),
      N = n_modelb_primary,
      model = paste0("Model B: Mixture*CKD interaction [", modelb_primary_label, "]")
    )

  # Design-based marginal effects via svycontrast with named contrast vectors (pooled across imputations if MI).

  # Robustly identify the interaction term regardless of factor coding.
  int_name <- cf_names[str_detect(cf_names, "mixture_qscore:ckd")]
  if (length(int_name) != 1) {
    warning("Could not uniquely identify mixture*ckd interaction term among: ",
            paste(cf_names, collapse = ", "), ". Skipping marginal effects.")
    tab3 <- tab3_raw
  } else {
    c_ckd0 <- setNames(rep(0, length(cf_names)), cf_names); c_ckd0["mixture_qscore"] <- 1
    c_ckd1 <- setNames(rep(0, length(cf_names)), cf_names); c_ckd1["mixture_qscore"] <- 1; c_ckd1[int_name] <- 1

    ctr0 <- contrast_pool(fit_int, c_ckd0)
    ctr1 <- contrast_pool(fit_int, c_ckd1)

    make_contrast_row <- function(ctr, label) {
      tibble(term = label, beta = fmt(ctr$estimate),
             conf.low = fmt(ctr$conf.low),
             conf.high = fmt(ctr$conf.high),
             p.value = fmt_p(ctr$p.value),
             N = n_modelb_primary, model = "Marginal effects (design-based)")
    }

    p_int_tidy <- int_tidy %>% filter(term == int_name) %>% pull(p.value)
    p_interaction <- if (length(p_int_tidy) > 0) p_int_tidy[1] else NA_real_

    tab3_marginal <- bind_rows(
      make_contrast_row(ctr0, "Mixture effect among CKD=0"),
      make_contrast_row(ctr1, "Mixture effect among CKD=1"),
      tibble(term = "P-interaction", beta = "", conf.low = "", conf.high = "",
             p.value = fmt_p(p_interaction), N = n_modelb_primary, model = "Marginal effects (design-based)")
    )
    tab3 <- bind_rows(tab3_raw, tab3_marginal)
  }
} else {
  tab3 <- tibble(
    term = c("mixture_qscore", "ckd", "mixture_qscore:ckd"),
    beta = "NA", conf.low = "NA", conf.high = "NA", p.value = "NA",
    N = n_modelb_primary, model = "Model B: Mixture*CKD interaction (not estimable)"
  )
}
kable(tab3, caption = "Table 3. Interaction model with design-based marginal effects by CKD group")
cat("\n*Table 3 note:* Model B is a survey-weighted linear model (`svyglm`) with interaction term `mixture_qscore * ckd`. ",
    "Inference: ", modelb_primary_label, " (N = ", format(n_modelb_primary, big.mark = ","),
    "); complete-case sensitivity N = ", format(n_modelb, big.mark = ","),
    if (mice_active) paste0("; MICE m = ", config$mice_m, " imputations of missing covariates (outcome/exposures/design not imputed). ") else ". ",
    "Intended covariates: ", paste(intended_covars_int, collapse = ", "), ". ",
    "Realized covariates (after complete-case support checks): ", paste(realized_covars_int, collapse = ", "), ". ",
    "Marginal effects for CKD=0 and CKD=1 are estimated using design-based contrasts (`svycontrast`); ",
    "for MI runs, contrasts are computed per imputation and pooled with Rubin's rules (mitools::MIcombine). ",
    "95% CI and p-values use t-distribution on Rubin (or survey residual) degrees of freedom.\n", sep = "")

# Table 4: sensitivity analyses -- using same covariate set as primary models.
sens_covars_rhs <- paste(realized_covars_int, collapse = " + ")
make_rhs_formula <- function(outcome, main_terms, covar_rhs) {
  rhs <- paste(main_terms, collapse = " + ")
  if (!is.null(covar_rhs) && nzchar(covar_rhs)) rhs <- paste(rhs, covar_rhs, sep = " + ")
  as.formula(paste(outcome, "~", rhs))
}
prep_svy_frame <- function(data_df) {
  data_svy <- as.data.frame(data_df)
  for (v in c("phenoage_accel", "phenoage_nokid", "phenoage_nokid_both", "phenoage_accel_unweighted",
              "mixture_qscore", "age", "pir", "bmi", "weight_combined", "psu", "strata")) {
    if (v %in% names(data_svy)) data_svy[[v]] <- suppressWarnings(as.numeric(as.character(data_svy[[v]])))
  }
  if ("ckd" %in% names(data_svy)) data_svy$ckd <- as.integer(data_svy$ckd)
  if ("ckd_sens" %in% names(data_svy)) data_svy$ckd_sens <- as.integer(data_svy$ckd_sens)
  for (v in c("sex", "race_eth", "cycle")) {
    if (v %in% names(data_svy)) data_svy[[v]] <- factor(as.character(data_svy[[v]]))
  }
  if ("smoker3"  %in% names(data_svy)) data_svy$smoker3  <- factor(as.character(data_svy$smoker3),  levels = c("Never", "Former", "Current"))
  if ("alcohol3" %in% names(data_svy)) data_svy$alcohol3 <- factor(as.character(data_svy$alcohol3), levels = c("Never", "Former", "Current"))
  data_svy$sex      <- droplevels(factor(data_svy$sex))
  data_svy$race_eth <- droplevels(factor(data_svy$race_eth))
  data_svy$cycle    <- droplevels(factor(data_svy$cycle))
  if ("smoker3"  %in% names(data_svy)) data_svy$smoker3  <- droplevels(data_svy$smoker3)
  if ("alcohol3" %in% names(data_svy)) data_svy$alcohol3 <- droplevels(data_svy$alcohol3)
  data_svy
}

# MI-aware sensitivity fit. `data_df` = complete-case subset for CC fit.
# `mi_filter_fn` (optional, used when mice_active) = function(df) -> filtered imputed df.
# Pooled MI inference is preferred over CC when available.
fit_svy_with_covar_guard <- function(data_df, outcome, main_terms, covars_terms, label,
                                     mi_filter_fn = NULL) {
  build_formula <- function(covs) {
    covar_rhs <- if (length(covs) > 0) paste(covs, collapse = " + ") else ""
    make_rhs_formula(outcome, main_terms, covar_rhs)
  }
  covs <- covars_terms
  intended <- covs
  f <- build_formula(covs)
  mf <- NULL
  repeat {
    mf <- tryCatch(model.frame(f, data = data_df, na.action = na.omit), error = function(e) NULL)
    if (is.null(mf) || nrow(mf) == 0) break
    factor_terms <- names(mf)[vapply(mf, is.factor, logical(1))]
    bad_factor_terms <- factor_terms[vapply(mf[factor_terms], function(x) nlevels(droplevels(x)) < 2, logical(1))]
    removable <- intersect(covs, bad_factor_terms)
    if (length(removable) == 0) break
    message("Sensitivity [", label, "]: dropping collapsed factor term(s): ", paste(removable, collapse = ", "))
    covs <- setdiff(covs, removable)
    f <- build_formula(covs)
  }
  if (is.null(mf) || nrow(mf) == 0) {
    return(list(
      fit = NULL, pooled = NULL, fits_mi = NULL,
      n_model = 0L, n_model_mi = NA_integer_, mi = FALSE,
      intended = intended, realized = covs,
      dropped = paste(setdiff(intended, covs), collapse = ", "),
      error = "No complete-case rows for model frame."
    ))
  }
  data_svy <- prep_svy_frame(data_df)
  des <- tryCatch(
    svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined, nest = TRUE, data = data_svy),
    error = function(e) svydesign(id = ~1, strata = NULL, weights = ~weight_combined, data = data_svy)
  )
  err <- NULL
  fit <- tryCatch(
    svyglm(f, design = des, family = gaussian()),
    error = function(e) { err <<- conditionMessage(e); NULL }
  )

  pooled <- NULL; fits_mi <- NULL; n_model_mi <- NA_integer_; err_mi <- NULL
  if (mice_active && !is.null(mi_filter_fn)) {
    mi_dfs <- tryCatch(
      lapply(seq_len(config$mice_m), function(i) {
        df <- imp_list$imputations[[i]]
        df <- mi_filter_fn(df)
        prep_svy_frame(df)
      }),
      error = function(e) { err_mi <<- conditionMessage(e); NULL }
    )
    if (!is.null(mi_dfs) && length(mi_dfs) > 0) {
      n_model_mi <- nrow(mi_dfs[[1]])
      mi_des <- tryCatch(
        survey::svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined, nest = TRUE,
                          data = mitools::imputationList(mi_dfs)),
        error = function(e) survey::svydesign(id = ~1, strata = NULL, weights = ~weight_combined,
                                              data = mitools::imputationList(mi_dfs))
      )
      pool_obj <- tryCatch(
        fit_svyglm_pool(f, mi_des),
        error = function(e) { err_mi <<- conditionMessage(e); NULL }
      )
      if (!is.null(pool_obj) && isTRUE(pool_obj$mi)) {
        pooled <- pool_obj$pooled
        fits_mi <- pool_obj$fits
      }
    }
  }
  list(
    fit = fit, pooled = pooled, fits_mi = fits_mi,
    n_model = nrow(mf), n_model_mi = n_model_mi,
    mi = !is.null(pooled),
    intended = intended, realized = covs,
    dropped = paste(setdiff(intended, covs), collapse = ", "),
    error = if (!is.null(pooled) && !is.null(err_mi)) paste(err, "; MI:", err_mi) else err
  )
}

# Wrap a sensitivity object as a `fit_svyglm_pool`-style object for downstream tidy/contrast helpers.
sens_as_fit <- function(obj) {
  if (isTRUE(obj$mi) && !is.null(obj$pooled) && !is.null(obj$fits_mi)) {
    list(fits = obj$fits_mi, pooled = obj$pooled, mi = TRUE)
  } else if (!is.null(obj$fit)) {
    list(fit = obj$fit, mi = FALSE)
  } else {
    NULL
  }
}

sens_ckd_egfr <- qdat %>% filter(!is.na(ckd_sens))
fit_sens1_obj <- fit_svy_with_covar_guard(
  sens_ckd_egfr, "phenoage_accel", c("mixture_qscore", "ckd_sens"), realized_covars_int,
  "CKD redefined as eGFR<60",
  mi_filter_fn = function(df) df[!is.na(df$ckd_sens), , drop = FALSE]
)
fit_sens1 <- fit_sens1_obj$fit

# PhenoAge-recomputed sensitivities use the precomputed columns on qdat / qdat_svy.
fit_sens2_obj <- fit_svy_with_covar_guard(
  qdat, "phenoage_nokid", c("mixture_qscore", "ckd"), realized_covars_int,
  "PhenoAge excluding creatinine only",
  mi_filter_fn = identity
)
fit_sens2 <- fit_sens2_obj$fit

fit_sens3_obj <- fit_svy_with_covar_guard(
  qdat, "phenoage_nokid_both", c("mixture_qscore", "ckd"), realized_covars_int,
  "PhenoAge excluding creatinine and albumin",
  mi_filter_fn = identity
)
fit_sens3 <- fit_sens3_obj$fit

get_mix_row <- function(obj, label, n_model_cc = NA_integer_, term_name = "mixture_qscore") {
  fit_for_tidy <- sens_as_fit(obj)
  if (is.null(fit_for_tidy)) {
    return(tibble(sensitivity = label, estimate = NA_real_, conf.low = NA_real_, conf.high = NA_real_,
                  p.value = NA_real_, N = n_model_cc, source = "Failed"))
  }
  tb <- tryCatch(tidy_pool(fit_for_tidy, conf.level = 0.95), error = function(e) NULL)
  if (is.null(tb) || !any(tb$term == term_name)) {
    return(tibble(sensitivity = label, estimate = NA_real_, conf.low = NA_real_, conf.high = NA_real_,
                  p.value = NA_real_, N = n_model_cc, source = "Failed"))
  }
  row <- tb %>% filter(term == term_name) %>% slice(1)
  N_used <- if (isTRUE(fit_for_tidy$mi)) (obj$n_model_mi %||% n_model_cc) else n_model_cc
  src    <- if (isTRUE(fit_for_tidy$mi)) "MI-pooled" else "Complete-case"
  tibble(sensitivity = label, estimate = row$estimate, conf.low = row$conf.low,
         conf.high = row$conf.high, p.value = row$p.value, N = N_used, source = src)
}
# Null-coalesce helper (R < 4.4 doesn't have %||%)
`%||%` <- function(a, b) if (is.null(a) || (is.atomic(a) && length(a) == 1 && is.na(a))) b else a

f_sens_uw <- make_rhs_formula("phenoage_accel_unweighted", c("mixture_qscore", "ckd"), sens_covars_rhs)
f_sens_both <- make_rhs_formula("phenoage_accel", c("mixture_qscore", "ckd"), sens_covars_rhs)
fit_sens_uw_obj <- fit_svy_with_covar_guard(
  qdat, "phenoage_accel_unweighted", c("mixture_qscore", "ckd"), realized_covars_int,
  "Unweighted residualized PhenoAgeAccel (sensitivity)",
  mi_filter_fn = identity
)
fit_sens_both_markers_obj <- fit_svy_with_covar_guard(
  qdat %>% filter(!is.na(egfr), !is.na(acr)), "phenoage_accel", c("mixture_qscore", "ckd"),
  realized_covars_int, "Restricted to participants with both eGFR and ACR",
  mi_filter_fn = function(df) df[!is.na(df$egfr) & !is.na(df$acr), , drop = FALSE]
)
fit_sens_no_acute_obj <- fit_svy_with_covar_guard(
  qdat %>% filter(is.na(crp_acute) | crp_acute == 0L), "phenoage_accel", c("mixture_qscore", "ckd"),
  realized_covars_int, "Exclude participants with CRP > 1 mg/dL",
  mi_filter_fn = function(df) df[is.na(df$crp_acute) | df$crp_acute == 0L, , drop = FALSE]
)
# Strict complete-case sensitivity stays CC by definition.
fit_sens_strict_obj <- fit_svy_with_covar_guard(
  qdat_strict, "phenoage_accel", c("mixture_qscore", "ckd"), realized_covars_int,
  "Strict complete-case (qgcomp-compatible PFAS subset), same primary adjustment as Model B",
  mi_filter_fn = NULL
)

# Sensitivity: additional adjustment for cardiometabolic conditions potentially on the causal pathway.
covars_cardiometabolic <- c(
  realized_covars_int,
  if (valid_num("diabetes", qdat)) "diabetes",
  if (valid_num("hypertension", qdat)) "hypertension",
  if (valid_num("cvd", qdat)) "cvd"
)
fit_sens_cardiometabolic_obj <- fit_svy_with_covar_guard(
  qdat, "phenoage_accel", c("mixture_qscore", "ckd"), covars_cardiometabolic,
  "Additional adjustment for diabetes, hypertension, CVD (potentially on causal pathway)",
  mi_filter_fn = identity
)
fit_sens_cardiometabolic <- fit_sens_cardiometabolic_obj$fit

tab4 <- bind_rows(
  get_mix_row(fit_sens1_obj, "CKD redefined as eGFR<60", fit_sens1_obj$n_model),
  get_mix_row(fit_sens2_obj, "PhenoAge excluding creatinine only", fit_sens2_obj$n_model),
  get_mix_row(fit_sens3_obj, "PhenoAge excluding creatinine and albumin", fit_sens3_obj$n_model),
  get_mix_row(fit_sens_uw_obj, "Unweighted residualized PhenoAgeAccel (sensitivity)", fit_sens_uw_obj$n_model),
  get_mix_row(fit_sens_both_markers_obj, "Restricted to participants with both eGFR and ACR", fit_sens_both_markers_obj$n_model),
  get_mix_row(fit_sens_no_acute_obj, "Exclude participants with CRP > 1 mg/dL", fit_sens_no_acute_obj$n_model),
  get_mix_row(fit_sens_cardiometabolic_obj, "Additional adjustment for diabetes, hypertension, CVD (potentially on causal pathway)", fit_sens_cardiometabolic_obj$n_model),
  get_mix_row(fit_sens_strict_obj, "Strict complete-case (qgcomp-compatible PFAS subset), same primary adjustment as Model B", fit_sens_strict_obj$n_model)
) %>%
  transmute(
    sensitivity,
    beta = fmt(estimate),
    conf.low = fmt(conf.low),
    conf.high = fmt(conf.high),
    p.value = fmt_p(p.value),
    N,
    source
  )
kable(tab4, caption = "Table 4. Sensitivity analyses (primary adjustment: sex, race, PIR, BMI, smoking [3-level], alcohol [3-level], cycle).")
cat("\n*Table 4 note:* Primary and most sensitivity models use the lean adjustment set (sex, race/ethnicity, PIR, BMI, smoking [Never/Former/Current], alcohol [Never/Former/Current], cycle). ",
    "Each row reports the `mixture_qscore` coefficient from a survey-weighted model (`svyglm`) with NHANES design variables. ",
    "The `source` column distinguishes MI-pooled (Rubin's rules across m=", config$mice_m,
    " imputations) versus complete-case rows; the strict-CC sensitivity remains complete-case by definition.\n", sep = "")

sens_diagnostics <- bind_rows(
  tibble(sensitivity = "CKD redefined as eGFR<60", N_model = fit_sens1_obj$n_model, dropped_covariates = ifelse(fit_sens1_obj$dropped == "", "None", fit_sens1_obj$dropped), fit_error = ifelse(is.null(fit_sens1_obj$error) || fit_sens1_obj$error == "", "None", fit_sens1_obj$error)),
  tibble(sensitivity = "PhenoAge excluding creatinine only", N_model = fit_sens2_obj$n_model, dropped_covariates = ifelse(fit_sens2_obj$dropped == "", "None", fit_sens2_obj$dropped), fit_error = ifelse(is.null(fit_sens2_obj$error) || fit_sens2_obj$error == "", "None", fit_sens2_obj$error)),
  tibble(sensitivity = "PhenoAge excluding creatinine and albumin", N_model = fit_sens3_obj$n_model, dropped_covariates = ifelse(fit_sens3_obj$dropped == "", "None", fit_sens3_obj$dropped), fit_error = ifelse(is.null(fit_sens3_obj$error) || fit_sens3_obj$error == "", "None", fit_sens3_obj$error)),
  tibble(sensitivity = "Unweighted residualized PhenoAgeAccel (sensitivity)", N_model = fit_sens_uw_obj$n_model, dropped_covariates = ifelse(fit_sens_uw_obj$dropped == "", "None", fit_sens_uw_obj$dropped), fit_error = ifelse(is.null(fit_sens_uw_obj$error) || fit_sens_uw_obj$error == "", "None", fit_sens_uw_obj$error)),
  tibble(sensitivity = "Restricted to participants with both eGFR and ACR", N_model = fit_sens_both_markers_obj$n_model, dropped_covariates = ifelse(fit_sens_both_markers_obj$dropped == "", "None", fit_sens_both_markers_obj$dropped), fit_error = ifelse(is.null(fit_sens_both_markers_obj$error) || fit_sens_both_markers_obj$error == "", "None", fit_sens_both_markers_obj$error)),
  tibble(sensitivity = "Exclude participants with CRP > 1 mg/dL", N_model = fit_sens_no_acute_obj$n_model, dropped_covariates = ifelse(fit_sens_no_acute_obj$dropped == "", "None", fit_sens_no_acute_obj$dropped), fit_error = ifelse(is.null(fit_sens_no_acute_obj$error) || fit_sens_no_acute_obj$error == "", "None", fit_sens_no_acute_obj$error)),
  tibble(sensitivity = "Additional adjustment for diabetes, hypertension, CVD (potentially on causal pathway)", N_model = fit_sens_cardiometabolic_obj$n_model, dropped_covariates = ifelse(fit_sens_cardiometabolic_obj$dropped == "", "None", fit_sens_cardiometabolic_obj$dropped), fit_error = ifelse(is.null(fit_sens_cardiometabolic_obj$error) || fit_sens_cardiometabolic_obj$error == "", "None", fit_sens_cardiometabolic_obj$error)),
  tibble(sensitivity = "Strict complete-case (qgcomp-compatible PFAS subset), same primary adjustment as Model B", N_model = fit_sens_strict_obj$n_model, dropped_covariates = ifelse(fit_sens_strict_obj$dropped == "", "None", fit_sens_strict_obj$dropped), fit_error = ifelse(is.null(fit_sens_strict_obj$error) || fit_sens_strict_obj$error == "", "None", fit_sens_strict_obj$error))
)
kable(sens_diagnostics, caption = "Sensitivity model diagnostics (dropped covariates and fit errors).")

# Supplementary: single-pollutant models with same primary adjustment as Model B.
single_pollutant_out <- lapply(exposure_vars, function(v) {
  sp_obj <- fit_svy_with_covar_guard(
    qdat, "phenoage_accel", c(v, "ckd"), realized_covars_int,
    paste0("Single pollutant: ", v),
    mi_filter_fn = function(df) df[!is.na(df[[v]]), , drop = FALSE]
  )
  diag_row <- tibble(
    exposure = v,
    N_model = if (isTRUE(sp_obj$mi)) sp_obj$n_model_mi else sp_obj$n_model,
    source = if (isTRUE(sp_obj$mi)) "MI-pooled" else "Complete-case",
    dropped_covariates = ifelse(sp_obj$dropped == "", "None", sp_obj$dropped),
    fit_error = ifelse(is.null(sp_obj$error) || sp_obj$error == "", "None", sp_obj$error)
  )
  row <- get_mix_row(sp_obj, paste0("Single pollutant: ", v),
                     n_model_cc = sp_obj$n_model, term_name = v)
  list(
    row  = tibble(
      exposure = v,
      beta = row$estimate, conf.low = row$conf.low, conf.high = row$conf.high,
      p.value = row$p.value, N = row$N, source = row$source
    ),
    diag = diag_row
  )
})
single_pollutant <- bind_rows(lapply(single_pollutant_out, `[[`, "row")) %>%
  mutate(
    beta = fmt(beta),
    conf.low = fmt(conf.low),
    conf.high = fmt(conf.high),
    p.value = fmt_p(p.value)
  )
single_pollutant_diagnostics <- bind_rows(lapply(single_pollutant_out, `[[`, "diag"))

kable(
  single_pollutant,
  caption = "Supplementary Table. Single-pollutant associations with PhenoAge acceleration (survey-weighted, same primary adjustment as Model B); MI-pooled when MICE is active."
)
kable(single_pollutant_diagnostics, caption = "Supplementary model diagnostics (single-pollutant N, source, dropped covariates, fit errors).")
```

# Figures

```{r figure}
dir.create("figures", showWarnings = FALSE)
n_chems <- length(primary_exposure_vars)

# Figure 1: survey-weighted predicted lines per CKD stratum.
# Mirror the proven standalone logic: filter NAs, fall back design->weights-only->lm,
# normalize predict() output (svyglm returns svystat; lm returns atomic).
fit_one_stratum <- function(sub_df) {
  needed <- c("phenoage_accel", "mixture_qscore", "weight_combined", "psu", "strata")
  sub_df <- sub_df[stats::complete.cases(sub_df[, needed, drop = FALSE]), , drop = FALSE]
  if (nrow(sub_df) == 0) return(NULL)
  fit <- tryCatch(
    survey::svyglm(phenoage_accel ~ mixture_qscore,
                   design = survey::svydesign(id = ~psu, strata = ~strata,
                                              weights = ~weight_combined, nest = TRUE,
                                              data = sub_df),
                   family = gaussian()),
    error = function(e) NULL)
  if (is.null(fit)) {
    fit <- tryCatch(
      survey::svyglm(phenoage_accel ~ mixture_qscore,
                     design = survey::svydesign(id = ~1, weights = ~weight_combined,
                                                data = sub_df),
                     family = gaussian()),
      error = function(e) NULL)
  }
  if (is.null(fit)) {
    fit <- tryCatch(stats::lm(phenoage_accel ~ mixture_qscore,
                              data = sub_df, weights = sub_df$weight_combined),
                    error = function(e) NULL)
  }
  fit
}

pred_lines <- tryCatch({
  full_score_range <- range(qdat$mixture_qscore, na.rm = TRUE)
  pdat_list <- lapply(c(0L, 1L), function(ck) {
    sub_df <- qdat %>% filter(ckd == ck)
    fit_pred <- fit_one_stratum(sub_df)
    if (is.null(fit_pred)) return(NULL)
    score_seq <- seq(full_score_range[1], full_score_range[2], length.out = 100)
    newdat <- data.frame(mixture_qscore = score_seq)
    pr <- tryCatch(
      predict(fit_pred, newdata = newdat, se.fit = TRUE, type = "response"),
      error = function(e) NULL)
    if (is.null(pr)) return(NULL)
    fit_v <- as.numeric(if (is.list(pr)) pr$fit else pr)
    se_v  <- if (is.list(pr) && !is.null(pr$se.fit)) as.numeric(pr$se.fit)
             else as.numeric(survey::SE(pr))
    newdat$fit <- fit_v
    newdat$se  <- se_v
    newdat$ckd <- ifelse(ck == 1L, "CKD", "No CKD")
    cf <- coef(fit_pred)
    vc <- tryCatch(stats::vcov(fit_pred), error = function(e) NULL)
    slope <- as.numeric(cf["mixture_qscore"])
    se_b  <- if (!is.null(vc)) sqrt(vc["mixture_qscore", "mixture_qscore"]) else NA_real_
    newdat$slope    <- slope
    newdat$slope_lo <- if (is.finite(se_b)) slope - 1.96 * se_b else NA_real_
    newdat$slope_hi <- if (is.finite(se_b)) slope + 1.96 * se_b else NA_real_
    newdat
  })
  pdat_list <- Filter(Negate(is.null), pdat_list)
  if (length(pdat_list) == 0) NULL else bind_rows(pdat_list)
}, error = function(e) NULL)

# Overall marginal direction: as mixture increases, what happens to phenoage_accel (pooled model).
overall_direction <- tryCatch({
  svy_full <- svydesign(id = ~psu, strata = ~strata, weights = ~weight_combined, nest = TRUE, data = qdat)
  fit_overall <- svyglm(phenoage_accel ~ mixture_qscore, design = svy_full, family = gaussian())
  score_seq <- seq(min(qdat$mixture_qscore, na.rm = TRUE), max(qdat$mixture_qscore, na.rm = TRUE), length.out = 100)
  newdat <- data.frame(mixture_qscore = score_seq)
  preds <- predict(fit_overall, newdata = newdat, se.fit = TRUE, type = "response")
  newdat$fit <- as.numeric(preds$fit)
  newdat
}, error = function(e) NULL)

# Faceted hex density of (mixture quantile score, PhenoAge acceleration) by CKD,
# with a thick survey-weighted trend line + 95% CI ribbon and per-facet slope label.
qdat_plot <- qdat %>%
  mutate(ckd_label = factor(ifelse(ckd == 0L, "No CKD", "CKD"),
                            levels = c("No CKD", "CKD")))
pred_plot <- if (!is.null(pred_lines)) {
  pred_lines %>%
    mutate(ckd_label = factor(ckd, levels = c("No CKD", "CKD")))
} else NULL

# Per-facet slope annotation: prefer MI-pooled marginal effects from fit_int
# (design-based contrasts already used in Table 3); fall back to per-stratum fits.
mi_ann <- tryCatch({
  if (exists("fit_int") && !is.null(fit_int)) {
    cf_names_a <- names(coef_pool(fit_int))
    int_name_a <- cf_names_a[stringr::str_detect(cf_names_a, "mixture_qscore:ckd")]
    if (length(int_name_a) == 1 && "mixture_qscore" %in% cf_names_a) {
      c0 <- setNames(rep(0, length(cf_names_a)), cf_names_a); c0["mixture_qscore"] <- 1
      c1 <- setNames(rep(0, length(cf_names_a)), cf_names_a); c1["mixture_qscore"] <- 1; c1[int_name_a] <- 1
      ctr0 <- contrast_pool(fit_int, c0)
      ctr1 <- contrast_pool(fit_int, c1)
      tibble::tibble(
        ckd_label = factor(c("No CKD", "CKD"), levels = c("No CKD", "CKD")),
        slope    = c(ctr0$estimate, ctr1$estimate),
        slope_lo = c(ctr0$conf.low, ctr1$conf.low),
        slope_hi = c(ctr0$conf.high, ctr1$conf.high)
      )
    } else NULL
  } else NULL
}, error = function(e) NULL)
ann_df <- if (!is.null(mi_ann)) mi_ann else if (!is.null(pred_plot)) {
  pred_plot %>%
    group_by(ckd_label) %>%
    summarise(slope = unique(slope), slope_lo = unique(slope_lo),
              slope_hi = unique(slope_hi), .groups = "drop")
} else NULL
if (!is.null(ann_df)) {
  ann_df <- ann_df %>%
    mutate(label = ifelse(
      is.finite(slope_lo) & is.finite(slope_hi),
      sprintf("beta = %+0.2f years per quantile increase (95%% CI %+0.2f, %+0.2f)",
              slope, slope_lo, slope_hi),
      sprintf("beta = %+0.2f years per quantile increase", slope)
    ))
}

theme_facet_box <- theme_bw(base_size = 11) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.4),
        strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        legend.position = "right",
        legend.key.height = unit(0.6, "lines"))

trend_color <- "#000000" # bright red, prints as a strong dark gray

fig1_hex <- ggplot(qdat_plot, aes(x = mixture_qscore, y = phenoage_accel)) +
  geom_hex(bins = 35) +
  scale_fill_gradient(low = "grey92", high = "black", name = "Count") +
  facet_wrap(~ ckd_label, ncol = 2) +
  scale_x_continuous(limits = range(qdat_plot$mixture_qscore, na.rm = TRUE),
                     expand = expansion(mult = c(0.01, 0.01))) +
  labs(x = "Mixture quantile score",
       y = "PhenoAge acceleration (years)") +
  theme_facet_box
if (!is.null(pred_plot)) {
  fig1_hex <- fig1_hex +
    geom_ribbon(data = pred_plot,
                aes(x = mixture_qscore,
                    ymin = fit - 1.96 * se, ymax = fit + 1.96 * se,
                    group = ckd_label),
                fill = trend_color, alpha = 0.20, inherit.aes = FALSE) +
    geom_line(data = pred_plot,
              aes(x = mixture_qscore, y = fit, group = ckd_label),
              color = trend_color, linewidth = 1.3, inherit.aes = FALSE)
}
if (!is.null(ann_df)) {
  y_top <- max(qdat_plot$phenoage_accel, na.rm = TRUE)
  x_left <- min(qdat_plot$mixture_qscore, na.rm = TRUE)
  fig1_hex <- fig1_hex +
    geom_text(data = ann_df,
              aes(x = x_left, y = y_top, label = label),
              hjust = 0, vjust = 1.2, size = 3.2, color = trend_color,
              fontface = "bold", inherit.aes = FALSE)
}
fig1_hex
ggsave("figures/fig1_hex_facet_CKD.jpg", plot = fig1_hex,
       width = 9.5, height = 5, dpi = 600, device = "jpeg",
       quality = 95, bg = "white")

# Figure 2: forest plot of marginal mixture effects (design-based).
fig2 <- NULL
if (!is.null(fit_int)) {
  cf_names <- names(coef_pool(fit_int))
  int_name <- cf_names[str_detect(cf_names, "mixture_qscore:ckd")]
  if (length(int_name) == 1 && "mixture_qscore" %in% cf_names) {
    c_ckd0 <- setNames(rep(0, length(cf_names)), cf_names); c_ckd0["mixture_qscore"] <- 1
    c_ckd1 <- setNames(rep(0, length(cf_names)), cf_names); c_ckd1["mixture_qscore"] <- 1; c_ckd1[int_name] <- 1
    ctr0 <- contrast_pool(fit_int, c_ckd0)
    ctr1 <- contrast_pool(fit_int, c_ckd1)

    mk_row <- function(ctr, grp) {
      tibble(
        group = grp,
        estimate = ctr$estimate,
        conf.low = ctr$conf.low,
        conf.high = ctr$conf.high,
        p.value = ctr$p.value
      )
    }
    forest_df <- bind_rows(
      mk_row(ctr0, "No CKD"),
      mk_row(ctr1, "CKD")
    ) %>%
      mutate(group = factor(group, levels = c("No CKD", "CKD")))

    fig2 <- ggplot(forest_df, aes(x = estimate, y = group, color = group)) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
      geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.18, linewidth = 0.8) +
      geom_point(size = 2.8) +
      geom_text(aes(label = ifelse(is.na(p.value), "",
                                   ifelse(p.value < 0.001, "p<0.001",
                                          paste0("p=", format(round(p.value, 3), nsmall = 3))))),
                nudge_y = 0.2, size = 3, color = "black", show.legend = FALSE) +
      scale_color_manual(values = c("No CKD" = "#1b9e77", "CKD" = "#d95f02")) +
      labs(
        title = "Marginal mixture effects by CKD status",
        subtitle = "Design-based contrasts from interaction model",
        x = "Beta (95% CI) for mixture score",
        y = NULL
      ) +
      theme_classic(base_size = 11) +
      theme(legend.position = "none")
    fig2
    ggsave("figures/fig2_forest_marginal_effects.jpg", plot = fig2, width = 7, height = 4, dpi = 300, device = "jpeg", quality = 95, bg = "white")
  }
}

# Figure 3: MICE convergence diagnostics in MCMC-style plate
# (Trace | Autocorrelation | Kernel density) per variable, boxed panels, no overall title.
fig3 <- NULL
if (exists("mids_obj") && !is.null(mids_obj)) {
  miss_n <- vapply(mice_target_vars, function(v) sum(is.na(qdat_svy[[v]])), integer(1))
  top_vars <- names(sort(miss_n, decreasing = TRUE))[seq_len(min(4, length(miss_n)))]
  top_vars <- top_vars[miss_n[top_vars] > 0]

  if (length(top_vars) >= 1) {
    # ---- Trace: chainMean is [variable, iteration, chain] ----
    cm <- mids_obj$chainMean
    trace_df <- as.data.frame.table(cm, responseName = "value", stringsAsFactors = FALSE) %>%
      dplyr::rename(variable = Var1, iter_lab = Var2, chain_lab = Var3) %>%
      dplyr::mutate(
        iter  = as.integer(sub("[^0-9]+", "", iter_lab)),
        chain = as.integer(sub("[^0-9]+", "", chain_lab))
      ) %>%
      dplyr::filter(variable %in% top_vars, !is.na(value))

    # ---- ACF: per chain, averaged across chains ----
    max_iter <- max(trace_df$iter, na.rm = TRUE)
    lag_max  <- max(1L, min(8L, max_iter - 1L))
    acf_ci   <- 1.96 / sqrt(max_iter) # white-noise 95% CI for ACF (Bartlett)
    acf_df <- trace_df %>%
      dplyr::group_by(variable, chain) %>%
      dplyr::arrange(iter, .by_group = TRUE) %>%
      dplyr::summarise(
        lag = list(0:lag_max),
        acf = list(as.numeric(stats::acf(value, plot = FALSE, lag.max = lag_max)$acf)),
        .groups = "drop"
      ) %>%
      tidyr::unnest(c(lag, acf)) %>%
      dplyr::group_by(variable, lag) %>%
      dplyr::summarise(acf = mean(acf, na.rm = TRUE), .groups = "drop")

    # ---- Density: observed (.imp == 0) vs imputed ----
    long_full <- mice::complete(mids_obj, "long", include = TRUE) %>%
      dplyr::mutate(source = ifelse(.imp == 0, "Observed", "Imputed"))
    var_pieces <- lapply(top_vars, function(v) {
      x <- long_full[[v]]
      val <- if (is.factor(x)) as.numeric(x) else suppressWarnings(as.numeric(x))
      data.frame(source = long_full$source, variable = v, value_num = val,
                 stringsAsFactors = FALSE)
    })
    long_imp <- do.call(rbind, var_pieces) %>%
      dplyr::filter(!is.na(value_num)) %>%
      dplyr::mutate(source = factor(source, levels = c("Observed", "Imputed")))

    # ---- ResearchGate-style theme: boxed, no grid, B&W ----
    bw_boxed <- function(base_size = 9) {
      theme_bw(base_size = base_size) +
        theme(
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, linewidth = 0.35),
          plot.background = element_rect(fill = "white", color = NA),
          plot.title = element_text(hjust = 0.5, face = "plain", size = base_size + 1),
          legend.position = "none"
        )
    }

    build_row <- function(var_name, col_titles) {
      tdf <- dplyr::filter(trace_df, variable == var_name)
      adf <- dplyr::filter(acf_df,   variable == var_name)
      ddf <- dplyr::filter(long_imp, variable == var_name)

      p_trace <- ggplot(tdf, aes(x = iter, y = value, group = chain)) +
        geom_line(alpha = 0.35, linewidth = 0.3, color = "black") +
        scale_x_continuous(expand = c(0.02, 0)) +
        labs(x = "Iteration", y = var_name,
             title = if (col_titles) "Traceplot" else NULL) +
        bw_boxed()

      p_acf <- ggplot(adf, aes(x = lag, y = acf)) +
        geom_hline(yintercept = 0, color = "black", linewidth = 0.35) +
        geom_hline(yintercept = c(-acf_ci, acf_ci), color = "gray55", linewidth = 0.25) +
        geom_segment(aes(xend = lag, yend = 0), linewidth = 0.35, color = "black") +
        geom_point(size = 1.6, color = "gray20") +
        scale_y_continuous(limits = c(-0.5, 1), expand = c(0, 0)) +
        labs(x = "Lag", y = "ACF",
             title = if (col_titles) "Autocorrelation" else NULL) +
        bw_boxed()

      # KDE: observed (solid) vs imputed (dashed); vertical lines at observed 2.5/97.5%
      q_obs <- stats::quantile(ddf$value_num[ddf$source == "Observed"],
                               probs = c(0.025, 0.975), na.rm = TRUE)
      p_kde <- ggplot(ddf, aes(x = value_num, linetype = source)) +
        geom_density(color = "black", linewidth = 0.45, na.rm = TRUE) +
        geom_vline(xintercept = q_obs, color = "gray45",
                   linewidth = 0.35, linetype = "solid") +
        scale_linetype_manual(values = c("Observed" = "solid", "Imputed" = "dashed")) +
        labs(x = "Value", y = "Density",
             title = if (col_titles) "Kernel density estimation" else NULL) +
        bw_boxed()

      p_trace + p_acf + p_kde + plot_layout(ncol = 3)
    }

    rows <- lapply(seq_along(top_vars), function(i) build_row(top_vars[i], col_titles = (i == 1L)))
    fig3 <- wrap_plots(rows, ncol = 1)

    fig3
    ggsave("figures/fig3_mice_convergence.jpg", plot = fig3,
           width = 8.2, height = 1.55 * length(top_vars) + 0.6,
           dpi = 300, device = "jpeg", quality = 95, bg = "white")
  }
}
```

# Cautions and Limitations

- Cross-sectional NHANES data do not establish temporality.
- PhenoAge includes kidney-related biomarkers (creatinine, albumin), creating potential circularity with CKD definitions. Sensitivity analyses recompute PhenoAge excluding (a) creatinine only and (b) both creatinine and albumin; both are non-standard derivatives and interpreted as robustness checks.
- PhenoAge uses Levine (2018) coefficients with correct units: albumin g/L (NHANES g/dL × 10), creatinine µmol/L (mg/dL × 88.4), glucose mmol/L (mg/dL ÷ 18.018). Model B uses phenoage_accel (age-residualized) and does not adjust for age again. Mixture quantiles are computed within each NHANES cycle to avoid temporal confounding (PFAS trends across retained cycles). Survey design uses full PSU/strata when possible; if the design fails (e.g. singleton PSUs), a weights-only fallback is used and noted.
- `qgcomp` point estimates are presented descriptively (pooled sampling weights applied, but PSU/strata not incorporated for variance estimation). Statistical inference relies on the survey-weighted regression models (`svyglm`) above. Bootstrap variance estimation incorporating replicate weights was not implemented because NHANES does not provide replicate weights for environmental subsamples across all cycles.
- LOD handling follows NHANES detection-flag conventions when indicator columns are available (`*LC` or `*L`): values flagged below LOD (indicator == 1) are replaced with value / sqrt(2). When no LOD indicator is available, values are left unchanged (no fallback NA imputation), which improves transparency but can reduce complete-case sample size.
- **Weight selection**: a weight audit table is printed and `config$weight_type = "auto"` applies the top-coverage eligible weight. For PFAS, auto-selection excludes full-sample interview/MEC weights (`WTINT2YR`, `WTMEC2YR`) and prioritizes environmental subsample weights (`WTS*2YR`), including the combined cross-cycle PFAS weight column.
- The primary PFAS mixture was defined as a fixed four-chemical set available across all retained cycles (PFOS, PFOA, PFNA, and PFHxS), with complete-case requirement on that set. PFDA was not consistently available across cycles and was therefore excluded from the primary mixture and evaluated only in supplementary analyses. Strict complete-case (qgcomp-compatible PFAS subset) modeling is reported as sensitivity. A sensitivity analysis using an available-exposure (adaptive) mixture can be added to assess robustness when using broader PFAS information.
- Exposure quantile cutpoints are defined within cycle using cycle-specific NHANES sampling weights.
- PhenoAgeAccel is computed as the residual from a survey-weighted regression of PhenoAge on chronological age (`svyglm`), ensuring the construct is population-representative. An unweighted residualization is included as sensitivity.
- CRP unit detection uses a two-part rule: (1) table-name provenance (HSCRP tables imply mg/L), and (2) a conservative distribution heuristic (median > 1.5 **and** p75 > 3). If either triggers, CRP is converted from mg/L to mg/dL (/ 10). Cycle-level diagnostic tables are printed. PhenoAge is recomputed after any conversions. Glucose is converted from mmol/L to mg/dL if median < 6.
- Extreme biomarker values are handled by plausibility filters (WBC > 50, glucose > 600, ALP > 500 set to missing). CRP > 1 mg/dL is retained in primary analyses and evaluated in sensitivity analyses excluding potential acute inflammatory states.
- Glucose is sourced from MEC biochemistry (`glu_mec`: BIOPRO/L40) for the primary analysis. The separate GLU table (fasting) is joined with renamed variables (`LBXGLU_GLU`, `LBXSGL_GLU`) to avoid duplicate columns; it is retained for diagnostics but not used in the primary model.
- Marginal effects in the interaction model (Table 3) use `svycontrast()` with design-based degrees of freedom and t-distribution p-values, maintaining full complex-survey consistency.
- Because complete-case requirements can collapse factor levels in the interaction model frame, Model B reports both intended and realized covariate sets; unsupported factor terms are dropped with explicit logging.
- **Multiple imputation (MICE).** When `config$use_mice = TRUE`, missing values on model covariates (smoking, alcohol, BMI, PIR, education, physical activity, diabetes, hypertension, CVD) are imputed within the Model B base sample (all participants with non-missing PhenoAge, CKD, design variables, and a complete fixed PFAS score). Outcome (PhenoAge / PhenoAgeAccel), exposures (PFAS, mixture quantile score), and survey design variables (PSU, strata, weights) are NEVER imputed. Imputation uses chained equations with `pmm` for continuous variables, `logreg` for binary factors, and `polyreg` for the 3-level Never/Former/Current factors. Model B and the survey-weighted sensitivity / single-pollutant analyses are then re-fit on each of the m imputed datasets and pooled with Rubin's rules via `mitools::MIcombine`; design-based marginal contrasts (`svycontrast`) are likewise computed per imputation and pooled. The complete-case Model B fit is retained and reported (in the Table 3 note) as a transparency check. **Table 2 (Model A overall mixture effect via `qgcomp`)** is also MI-pooled: `qgcomp.noboot` is fit on each imputed dataset and the overall psi (and its variance) are combined with Rubin's rules; a complete-case `qgcomp` row is reported alongside as a sensitivity check. `qgcomp` variance does not incorporate PSU/strata, so survey-design inference for the interaction question still relies on Model B (Table 3).

# Reproducibility

```{r session-info}
sessionInfo()
```
